### 2017-11-06 Analysis/Norm_concept_kids .Rmd
final pre registration for kids

---
title: "WordRec Experiment"
author: "A. Fourtassi & M. Frank"
date: "January 13, 2017"
output:
  html_document:
    number_sections: yes
    toc: yes
---

Libraries.

```{r}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(langcog)
theme_set(theme_bw())
```

Data.

```{r}

#Skip the pre task trials 
#Filter subjects who had no technical problem
#Select subjects who got more than 50% correct answers on the pre task trials

d_norm <- read_delim("Norm_kids.txt", delim = " ") %>%
  filter(type == "Train") #%>%
  #filter(score > 0.5)

```

Process norming data

```{r}
concept <- d_norm %>%
  filter(condition == 'concept') %>%
  group_by(distance) %>%
  multi_boot_standard(col = "answer") #%>%
  #rename(concept_dist = conceptOrd,
      #   semDist=mean)

```
Norming plots

```{r}
plot_concept  <- ggplot(concept, aes(x = distance, y= mean)) +
  geom_pointrange(aes(ymin = summary_ci_lower, ymax = summary_ci_upper), 
                  position = position_dodge(width = .03), size=0.5) +

 xlab("distance") +ylab("% different")+ theme_bw()

plot_concept

```