### 2018-08-13 Analysis/process.Rmd
analysis

---
title: "WordRec Experiment"
date: "January 13, 2017"
output:
  html_document:
    number_sections: yes
    toc: yes
---

Libraries.

```{r}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(langcog)
theme_set(theme_bw())
```

Data.

```{r}

d_word_summer <- read_delim("data_summer_2.txt", delim = " ") %>%
  filter(type == "Test") %>%
  filter(age != 6) %>%
  filter(code != 'test',
         !is.na(code)) %>%
  filter(grepl('soph', code))

d_all <- d_word_summer %>%
  mutate(iscorrect=ifelse(answer==correct, 1, 0))

```

Process data:

```{r}

data <- d_all %>%
  group_by(ID, sound_dist, concept_dist) %>%
  dplyr::summarise(mean = mean(iscorrect))
  #multi_boot_standard(col = "iscorrect")

#Catch
data_excl <- data %>%
  filter(sound_dist == 'catch') %>%
  ungroup() %>%
  select(ID, mean) %>%
  dplyr::rename(score = mean)

data_catch <- d_all %>%
  left_join(data_excl) %>%
  filter(score > 0.5) %>%
  filter(sound_dist != 'catch')

#Exclude items 
#item_excl <- (data %>%
#  filter(sound_dist != 'catch') %>%
#  filter(mean < 0.25) %>%
#  ungroup() %>%
#  distinct(ID))$ID

#data_reverse <- d_word %>%
# filter(!(ID %in% item_excl))

#data_catch_reverse <- d_word %>%
#  left_join(data_excl) %>%
#  filter(score >= 0.5) %>%
#  filter(!(ID %in% item_excl))

#data_final <- data_good %>%
#  mutate(age2 = ifelse(grepl('4',age), 4, 5))

```

Summary 
```{r }
#{r fig.width=2, fig.height=5}

sum_catch <- data_catch %>%
  group_by(sound_dist, concept_dist) %>%
  #group_by(sound_dist) %>%
  dplyr::summarise(mean = mean(iscorrect),
                   sd = sd(iscorrect),
                   n = n()) %>%
  mutate(se = sd / sqrt(n),
         lower = mean - qt(1 - (0.05 / 2), n - 1) * se,
         upper = mean + qt(1 - (0.05 / 2), n - 1) * se) %>%
  select(-sd, -n, -se)

ggplot(sum_catch, 
      aes(x = sound_dist, y = mean, col = factor(concept_dist))) +
  #geom_point()+
  geom_pointrange(aes(ymin = lower, ymax = upper), 
                 position = position_dodge(width = .1)) + 
  #geom_line(lty = 2) + 
  geom_hline(yintercept = 0.50) +
  xlab("Aud distance") +ylab("% Accurate")+
  scale_colour_discrete(name="Vis Dist")+
theme(legend.title = element_text(size=8))

#Model
data_model <- data_catch %>%
  mutate(s_dist=ifelse(sound_dist=="similar", 0, 1),
         c_dist=ifelse(concept_dist=="similar", 0, 1))

model <- glmer(iscorrect ~ s_dist + c_dist + 
                 (1 | ID) + 
                 (1 | item_object) + 
                 (1 | item_sound), 
               data = data_model, family = binomial())

summary(model)

```