### 2018-12-28 Analysis/process.Rmd
new modif on manuscript

---
title: "WordRec Experiment"
date: "January 13, 2017"
output:
  html_document:
    number_sections: yes
    toc: yes
---

Libraries.

```{r}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(langcog)
library(ggthemes)
library(lme4)
library(purrr)
theme_set(theme_few())
```

Adult data.

```{r}

d_adult <- read_delim("adult_pilot_3.txt", delim = " ")  %>%
  filter(type == "Test") %>%
  mutate(iscorrect=ifelse(answer==correct, 1, 0))

d_adult_score <- d_adult %>%
  group_by(ID, sound_dist, concept_dist) %>%
  dplyr::summarise(mean = mean(iscorrect))

d_adult_excl <- d_adult_score %>%
  filter(sound_dist == 'catch') %>%
  ungroup() %>%
  select(ID, mean) %>%
  dplyr::rename(score = mean)
  #select(ID, score)

d_adult_good <- d_adult %>%
  left_join(d_adult_excl) %>%
  filter(score >0.7) %>%
  filter(sound_dist != 'catch')

feather::write_feather(d_adult_good, "../saved_data/data_adult.feather")
```

Children data

```{r}
d_child <- read_delim("data_summer_all.txt", delim = " ") %>%
  filter(type == "Test") %>%
  filter(age != 6) %>%
  filter(code != 'test',
         !is.na(code)) %>%
  filter(grepl('soph', code))%>%
  mutate(iscorrect=ifelse(answer==correct, 1, 0))

d_child_score <- d_child %>%
  group_by(ID, sound_dist, concept_dist) %>%
  dplyr::summarise(mean = mean(iscorrect))

#Catch
d_child_excl <- d_child_score %>%
  filter(sound_dist == 'catch') %>%
  ungroup() %>%
  select(ID, mean) %>%
  dplyr::rename(score = mean)

d_child_good <- d_child %>%
  left_join(d_child_excl) %>%
  filter(score > 0.5) %>%
  filter(sound_dist != 'catch')

feather::write_feather(d_child_good, "../saved_data/data_child.feather")

```

Similarity data

```{r}

d_sim <- read_delim("sim_results.txt", delim = " ")  %>%
  filter(condition != "obj_prac" & condition != "sound_prac") %>%
  filter(problem == 'No') %>%
  mutate(scaled = (answer-1)/6)

d_sim_summary <- d_sim %>%
  group_by(condition, distance) %>%
  summarise(mean = mean(scaled),
                   sd = sd(scaled),
                   n = n()) %>%
  mutate(se = sd / sqrt(n),
         lower = mean - qt(1 - (0.05 / 2), n - 1) * se,
         upper = mean + qt(1 - (0.05 / 2), n - 1) * se) %>%
  select(-sd, -n, -se)

d_sim_transform <-  d_sim_summary %>%
  mutate(normalized =ifelse(condition=='sound_real', mean/0.926, mean)) %>% #Here I only normalize sounds, objects are already normalized
  filter(distance != "iden") %>%
  mutate(mydist = ifelse(distance=="0.6", 1, myscale))%>% # Here I create a ligne for different which were not explicitly asked from participants because it was used as a prototype, thus dist=1
  mutate(labels = ifelse(distance=='0.1','similar', 
                         ifelse(distance=='0.3','intermediate', 
                                ifelse(distance=='sim','similar',
                                       ifelse(distance=='med','intermediate', 'different')))))

#d_sim_summary$distance <- factor(sum_child$concept_dist, levels = c("similar", "different"))
#sum_child$sound_dist <- factor(sum_child$sound_dist, levels = c("similar", "different"))

ggplot(d_sim_transform, 
      aes(x = labels, y = mydist, col=condition)) +
  geom_point()+
  #geom_pointrange(aes(ymin = lower, ymax = upper), 
     #            position = position_dodge(width = .1)) + 
  xlab("Distance") +ylab("rating")+
  theme_few()+
theme(legend.title = element_text(size=11),
      legend.text=element_text(size=11),
      axis.text = element_text(size = 11)
      ) +
  ylim(c(0,1.1))

```

Summary

```{r}

sum_adult <- d_adult_good %>%
  group_by(sound_dist, concept_dist) %>%
  dplyr::summarise(mean = mean(iscorrect),
                   sd = sd(iscorrect),
                   n = n()) %>%
  mutate(se = sd / sqrt(n),
         lower.ci = mean - qt(1 - (0.05 / 2), n - 1) * se,
         upper.ci = mean + qt(1 - (0.05 / 2), n - 1) * se) %>%
  select(-sd, -n, -se)

sum_child <- d_child_good %>%
  group_by(sound_dist, concept_dist) %>%
  dplyr::summarise(mean = mean(iscorrect),
                   sd = sd(iscorrect),
                   n = n()) %>%
  mutate(se = sd / sqrt(n),
         lower = mean - qt(1 - (0.05 / 2), n - 1) * se,
         upper = mean + qt(1 - (0.05 / 2), n - 1) * se) %>%
  select(-sd, -n, -se)

```

Plots
```{r fig.width=3, fig.height=1.7}

sum_child$concept_dist <- factor(sum_child$concept_dist, levels = c("similar", "different"))
sum_child$sound_dist <- factor(sum_child$sound_dist, levels = c("similar", "different"))

plot_child <- ggplot(sum_child, 
      aes(x = sound_dist, y = mean, group = factor(concept_dist), col = factor(concept_dist))) +
  #geom_point()+
  geom_pointrange(aes(ymin = lower, ymax = upper), 
                 position = position_dodge(width = .1)) + 
  geom_line() + 
  geom_hline(yintercept = 0.50, linetype="dashed") +
  xlab("Auditory similary") +ylab("Accuracy")+
  scale_colour_discrete(name="Visual similarity")+
  theme_few()+
theme(legend.title = element_text(size=11),
      legend.text=element_text(size=11),
      axis.text = element_text(size = 11)
      ) #+ facet_grid(.~block)

plot_child

sum_adult$concept_dist <- factor(sum_adult$concept_dist, levels = c("similar", "different"))
sum_adult$sound_dist <- factor(sum_adult$sound_dist, levels = c("similar", "different"))

plot_adult <- ggplot(sum_adult, 
       aes(x = sound_dist, y = mean, group = factor(concept_dist), col = factor(concept_dist))) +

  geom_pointrange(aes(ymin = lower.ci, ymax = upper.ci), 
                  position = position_dodge(width = .1)) +

  geom_hline(yintercept = .50, linetype="dashed") + 
  geom_line() +
  xlab("Auditory similary") +ylab("Accuracy")+
  #scale_colour_manual(values = c("red", "blue"), name="Vis Dist")+
  scale_colour_discrete(name="Visual similarity")+

  #scale_y_continuous(limits = c(0.45, 0.9))+
  theme_few()+
theme(legend.title = element_text(size=11),
      legend.text=element_text(size=11),
      axis.text = element_text(size = 11)
      ) #+ facet_grid(. ~ block)+

#theme(plot.title = element_text(size = 12, face = "bold"),
#    legend.title=element_text(size=10), 
#    legend.text=element_text(size=9))
plot_adult

```

Models
```{r}

#Code and center levels
data_ch <- d_child_good %>%
  mutate(s_dist=ifelse(sound_dist=="similar", -1,1),
         c_dist=ifelse(concept_dist=="similar", -1, 1))

data_ad <- d_adult_good %>%
  mutate(s_dist=ifelse(sound_dist=="similar", -1,1),
         c_dist=ifelse(concept_dist=="similar", -1, 1))

#Maximal random structure
model_ch <- glmer(iscorrect ~ s_dist * c_dist + 
                 (s_dist * c_dist  | ID) + 
                 (c_dist | item_object) + 
                 (s_dist | item_sound), 
               data = data_ch, family = binomial(),
               glmerControl(optimizer="bobyqa", optCtrl = list(maxfun = 100000)))

model_ad <- glmer(iscorrect ~ s_dist * c_dist + 
                 (s_dist * c_dist  | ID) + 
                 (c_dist | sound) + 
                 (s_dist | concept), 
               data = data_ad, family = binomial(),
               glmerControl(optimizer="bobyqa", optCtrl = list(maxfun = 100000)))

```

Use values of perceptual distance from the similarity judgement task

```{r}
#Process the data to prepare for the fit
data_fit_adult <- sum_adult %>%
  rename(lower  = lower.ci,
         upper = upper.ci) %>%
  mutate(group = 'adults',
         s_dist = ifelse(sound_dist =='different',  0.66, 0.23),
         c_dist = ifelse(concept_dist =='different', 0.40, 0.19))

data_fit_child <- sum_child %>% 
   mutate(group = 'children',
         s_dist = ifelse(sound_dist =='different',  1, 0.66),
         c_dist = ifelse(concept_dist =='different', 1, 0.40))

data_fit <- data_fit_adult %>%
  bind_rows(data_fit_child)
```

Now I need to predict accuracy as a function of sound and meaning similarity 
```{r}

#This the analytical solution of the posterior + luce rule
combine_cues <- function(s, o, Vr) {
  s1 = 0
  o1 = 0
  no_acc <- exp(-(s1-s)^2/(2*(Vr)^2)) + exp(-(o1-o)^2/(2*(Vr)^2))
  acc <- 1 + exp(-((s1-s)^2+(o1-o)^2)/(2*(Vr)^2)) 
  acc_norm <- acc/(acc+no_acc)

  return(acc_norm)
}

combine_cues_2 <- function(s, o, Vr_s, Vr_o) {
  s1 = 0
  o1 = 0
  no_acc <- exp(-(s1-s)^2/(2*(Vr_s)^2)) + exp(-(o1-o)^2/(2*(Vr_o)^2))
  acc <- 1 + exp(-((s1-s)^2/(2*(Vr_s)^2)+(o1-o)^2/(2*(Vr_o)^2))) 
  acc_norm <- acc/(acc+no_acc)

  return(acc_norm)
}

fit_adults <- nls(mean ~ combine_cues(s_dist, c_dist, Vr), data=subset(data_fit, group=='adults'), start=list(Vr=0.1), lower=c(0), algorithm="port")

fit_adults_2 <- nls(mean ~ combine_cues_2(s_dist, c_dist, Vr_s, Vr_o), data=subset(data_fit, group=='adults'), start=list(Vr_s=0.1, Vr_o=0.04), lower=c(0,0), algorithm="port")

Vr_ad <- coef(fit_adults)
Vr_s_ad <- coef(fit_adults_2)[1]
Vr_o_ad <- coef(fit_adults_2)[2]

fit_children <- nls(mean ~ combine_cues(s_dist, c_dist, Vr), data=subset(data_fit, group=='children'), start=list(Vr=0.5), lower=c(0), algorithm="port")

fit_children_2 <- nls(mean ~ combine_cues_2(s_dist, c_dist, Vr_s, Vr_o), data=subset(data_fit, group=='children'), start=list(Vr_s=0.5, Vr_o=0.3), lower=c(0,0), algorithm="port")

Vr_ch <- coef(fit_children)
Vr_s_ch <- coef(fit_children_2)[1]
Vr_o_ch <- coef(fit_children_2)[2]

model1_val <- as.data.frame(confint2(fit_children)) %>%
  mutate(val= Vr_ch, 
         model=1, 
         group = "children") %>%
  bind_rows(as.data.frame(confint2(fit_adults)) %>%
              mutate(val= Vr_ad, 
                     model=1, 
                     group = "adults"))

model2_val <- add_rownames(as.data.frame(confint2(fit_children_2)), "param") %>%
  mutate(val= ifelse(param=='Vr_s',Vr_s_ch, Vr_o_ch),
         model=2, 
         group = "children") %>%
  bind_rows(add_rownames(as.data.frame(confint2(fit_adults_2)), "param") %>%
              mutate(val= ifelse(param=='Vr_s',Vr_s_ad, Vr_o_ad),
                     model=2, 
                     group = "adults"))

feather::write_feather(model1_val, "../saved_data/values_mod1.feather")       
feather::write_feather(model2_val, "../saved_data/values_mod2.feather")

model2_val <- c(confint(fit_children_2), Vr_s_ch, Vr_o_ch, model="2", group = "children") %>%
  bind_rows(c(confint(fit_adults_2), Vr_s_ad, Vr_o_ad, model="2", group = "adults"))

model_data <- data_fit %>%
  mutate(prediction = ifelse(group=='adults', 
                             combine_cues(s_dist, c_dist, Vr_ad),
                             combine_cues(s_dist, c_dist, Vr_ch)
                             )) %>%
  mutate(prediction_2 = ifelse(group=='adults', 
                             combine_cues_2(s_dist, c_dist, Vr_s_ad, Vr_o_ad),
                             combine_cues_2(s_dist, c_dist, Vr_s_ch, Vr_o_ch)
                             )) %>%
  rename(human = mean)

feather::write_feather(model_data, "../saved_data/data_correlations.feather")

model_data_plot <-  model_data %>%
  gather(key, value, human, prediction, prediction_2) %>%
  mutate(ci.upper = ifelse(key=='human', upper, value),
         ci.lower = ifelse(key=='human', lower, value))

feather::write_feather(model_data_plot, "../saved_data/data_human_model.feather")

```

Simulations
```{r}

simulation1 <- function(s,o,Vr) {
  s1 = 0
  o1 = 0
  no_acc <- exp(-(s1-s)^2/(2*Vr)) + exp(-(o1-o)^2/(2*Vr))
  acc <- 1 + exp(-((s1-s)^2+(o1-o)^2)/(2*Vr)) 
  acc_norm <- acc/(acc+no_acc)

  return(acc_norm)
}

s <- seq(0,1,0.01)

simulation_data <- data.frame(x=s, y=simulation1(s, o=0.25, Vr=(0.15)^2), sigma = '0.15', object_distance = '0.25') %>%
  bind_rows(data.frame(x=s, y=simulation1(s, o=0.5, Vr=(0.15)^2), sigma = '0.15', object_distance = '0.5')) %>%
  bind_rows(data.frame(x=s, y=simulation1(s, o=1, Vr=(0.15)^2), sigma = '0.15', object_distance = '1')) %>%
  bind_rows(data.frame(x=s, y=simulation1(s, o=0.25, Vr=(0.3)^2), sigma = '0.3', object_distance = '0.25')) %>%
  bind_rows(data.frame(x=s, y=simulation1(s, o=0.5, Vr=(0.3)^2), sigma = '0.3', object_distance = '0.5')) %>%
  bind_rows(data.frame(x=s, y=simulation1(s, o=1, Vr=(0.3)^2), sigma = '0.3', object_distance = '1')) %>%
  bind_rows(data.frame(x=s, y=simulation1(s, o=0.25, Vr=(0.6)^2), sigma = '0.6', object_distance = '0.25')) %>%
  bind_rows(data.frame(x=s, y=simulation1(s, o=0.5, Vr=(0.6)^2), sigma = '0.6', object_distance = '0.5')) %>%
  bind_rows(data.frame(x=s, y=simulation1(s, o=1, Vr=(0.6)^2), sigma = '0.6', object_distance = '1'))

feather::write_feather(simulation_data, "../saved_data/simulations.feather")

ggplot(simulation_data, aes(x=x, y=y, col=sigma)) + 
  geom_line() + 
  theme_few()+
  theme(legend.title = element_text(size=11),
      legend.text=element_text(size=11),
      axis.text = element_text(size = 11),
      aspect.ratio = 0.7
      )+
  ylim(c(0.4,1))+
  geom_hline(yintercept = 0.50, linetype="dashed") +
  xlab("sound distance") + ylab("Prob. accurate") +
  facet_grid(.~object_distance)

```

```{r}

model_data_plot$group <- factor(model_data_plot$group, levels = c("children", "adults"))

ggplot(model_data_plot, 
#ggplot(subset(model_data_plot, group=='children'),
      aes(x = sound_dist, y = value, group = factor(concept_dist), col = factor(concept_dist))) +
  #geom_point()+
  geom_pointrange(aes(ymin = ci.lower, ymax = ci.upper), 
                 position = position_dodge(width = .1),
                 size=0.2) + 
  geom_line() + 
  geom_hline(yintercept = 0.50, linetype="dashed") +
  xlab("Auditory similary") +ylab("Accuracy")+
  scale_colour_discrete(name="Visual similarity")+
  theme_few()+
theme(legend.title = element_text(size=11),
      legend.text=element_text(size=11),
      axis.text = element_text(size = 11)
      ) + facet_grid(group~key, scales="free_y") +
  theme(aspect.ratio = 0.7)

```

plots 
```{r}

Gdist <- function(s0, Vr, x){

  return((1/sqrt(2*pi*Vr))*exp(-(s0-x)^2/(2*Vr)))
}

s = seq(-1,2.5,0.01)
y = Gdist(0,1,x)

illustration <- data.frame(x=s, y=Gdist(s0=0, Vr=0.5, s), rep = "1", case = "low accuracy", type='Disambiguation') %>%
  bind_rows(data.frame(x=s, y=Gdist(s0=0.5, Vr=0.5, s), rep = "2", case = "low accuracy", type='Disambiguation')) %>%
  bind_rows(data.frame(x=s, y=Gdist(s0=0, Vr=0.5, s), rep = "1", case = "high accuracy", type='Disambiguation')) %>%
  bind_rows(data.frame(x=s, y=Gdist(s0=1.1, Vr=0.5, s), rep = "2", case = "high accuracy", type='Disambiguation')) %>%
  bind_rows(data.frame(x=s, y=Gdist(s0=0, Vr=0.5, s), rep = "1", case = "low accuracy", type='Development')) %>%
  bind_rows(data.frame(x=s, y=Gdist(s0=0.5, Vr=0.5, s), rep = "2", case = "low accuracy", type='Development')) %>%
  bind_rows(data.frame(x=s, y=Gdist(s0=0, Vr=0.1, s), rep = "1", case = "high accuracy", type='Development')) %>%
  bind_rows(data.frame(x=s, y=Gdist(s0=0.5, Vr=0.1, s), rep = "2", case = "high accuracy", type='Development'))

illustration$case <- factor(illustration$case, levels = c("low accuracy", "high accuracy"))
illustration$type <- factor(illustration$type, levels = c("Disambiguation", "Development"))

probs <- subset(subset(illustration, x==0)) %>%
  group_by(type, case) %>%
  mutate(prob = round(y/sum(y),2)) %>%
  filter(rep=='1')

ggplot(illustration, aes(x=x, y=y, linetype=rep)) + 
  geom_line() + 
  theme_few()+
  theme(legend.title = element_text(size=11),
      legend.text=element_text(size=11),
      axis.text = element_text(size = 11),
      strip.text = element_text(size = 11),
      aspect.ratio = 0.5
      )+
  #xlim(c(-1,1))+
  #ylim(c(0,1))+
  #geom_vline(xintercept = 0.0, color="red") +
  #geom_segment(aes(x = 0, y = 0, xend = 0, yend = y, colour = rep),
  #             position=position_dodge(width = .2), 
   #            data = subset(illustration, x==0))+

  geom_col(aes(x = x, y = y, fill = rep),
               position=position_dodge(), 
               width=0.2,
               data = subset(illustration, x==0))+
  geom_text(aes(label=paste("Acc=", prob, sep="")),
            x=2, y=1, size=4, 
            fontface = "bold",
            data=probs)+
  xlab("Perceptual space") + ylab("Prob") +
  facet_grid(type~case) +
  theme(legend.position="none")

```