### 2019-01-09 cogsci/cogsci.tex

% Template for Cogsci submission with R Markdown

% Stuff changed from original Markdown PLOS Template
\documentclass[10pt, letterpaper]{article}

\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{float}
\usepackage{caption}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}

% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% hyperref package, useful for hyperlinks
\usepackage{hyperref}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% Sweave(-like)
\usepackage{fancyvrb}
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
\newenvironment{Schunk}{}{}
\DefineVerbatimEnvironment{Code}{Verbatim}{}
\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
\newenvironment{CodeChunk}{}{}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace}
%\doublespacing

% % Text layout
% \topmargin 0.0cm
% \oddsidemargin 0.5cm
% \evensidemargin 0.5cm
% \textwidth 16cm
% \textheight 21cm

\title{Continuous Development in Learning Similar-sounding Words}

\usepackage{tipa}

\author{{\large \bf Abdellah Fourtassi} \\ \texttt{afourtas@stanford.edu} \\ Department of Psychology \\ Stanford University \And {\large \bf Sophie Regan} \\ \texttt{sregan20@stanford.edu} \\ Department of Psychology \\ Stanford University \And {\large \bf Michael C. Frank} \\ \texttt{mcfrank@stanford.edu} \\ Department of Psychology \\ Stanford University}

\begin{document}

\maketitle

\begin{abstract}
Stager and Werker (1997) first showed that children's early ability to
distinguish similar sounds do not automatically translate into word
learning skills. Early explanations proposed that children may not be
able to encode subtle phonetic contrasts when learning novel word
meanings, thus suggesting a discontinuous/stage-like pattern of
development. However, later work has revealed (e.g., through using
simpler testing methods) that children do encode such contrasts, thus
favoring a rather a continuous pattern of development. The present study
proposes a precise probabilistic model describing how development may
proceed in a continuous fashion across the lifespan. The purpose of the
model is to account for previously documented facts while providing new
predictions. We collected data from both preschool children and adults,
and we showed that the model can explain various patterns of learning
both within the same age and across development. Our work highlights the
role of computational modeling in advancing our understanding of
development both through organizing exsinting knowledge and through
generating new principled hypotheses.

\textbf{Keywords:}
word learning, cognitive development, computational modeling
\end{abstract}

\section{Introduction}\label{introduction}

Over the first year of life, children become sensitive to the phonetic
variations that are used to distinguish meanings in their native
language (Werker \& Tees, 1984). One could imagine that these perceptual
skills would be automatically applied to the task of word learning.
However, developmental data show that 14 m.o children find it
challenging to associate minimally different (but perceptually
discriminated) sounds such as ``bin'' and ``din'' to different objects
(Stager \& Werker, 1997).

Several factors can explain this finding. For example it is possible
that the task of meaning learning increases cognitive demands on
children (compared to a simple perceptual discrimination). In
particular, it requires paying attention to both the sounds and the
corresponding objects, which may hinder precise encoding in memory of
some phonetic details (Hofer \& Levy, 2017; Stager \& Werker, 1997).
Additional difficulty might arise from ambiguous phonological boundaries
at this stage of development (e.g., Rost \& McMurray, 2009), or from
uncertainty about the referential status of the novel word (Fennell \&
Waxman, 2010).

Regardless of the exact explanation, it is generally accepted that by
around 17 m.o, children succeed in the same task (Werker, Fennell,
Corcoran, \& Stager, 2002). What could be the mechanism of development?
Early accounts assumed--whether implicitly or explicitly-- that children
encode words in a binary way: they either fail or succeed in encoding
the relevant phonetic details (simultaneously with the meanings). This
account suggested a discontinuous/stage-like pattern of development
whereby younger children fail to encode the contrastive phonetic detail,
whereas older children succeed.

Nevertheless, subsequent findings suggest otherwise. On the one hand,
14-month-olds--who typically fail in the original task--succeed when an
easier \emph{testing} method is used, even under the same
\emph{learning} conditions (Yoshida, Fennell, Swingley, \& Werker,
2009). They also succeed when uncertainty is mitigated via
disambiguating cues (e.g., Thiessen, 2007). On the other hand, adults
show patterns of learning similar to those shown by 14-month-olds when
the task is more challenging and when similarity between words increases
(Pajak, Creel, \& Levy, 2016; White, Yee, Blumstein, \& Morgan, 2013).

These evidence point towards another scenario, where the representations
are encoded in a probabilistic (rather than binary) way, and where
development is continuous, rather than stage-like (Pajak et al., 2016;
Swingley, 2007; Yoshida et al., 2009). On this account, correct
representations are learned early in development, but these
representations are encoded with higher uncertainty in younger children,
leading to apparent failure in relatively demanding tasks. Development
is a continuous process whereby the initial noisy representations become
more precise. Crucially, more precise representation are still
imperfect: Even with adults show low accuracy learning when the sounds
are subtle enough, e.g., non-native sounds (Pajak et al., 2016).

We provide an intuitive illustration of how such an account explains
patterns of learning and development in Figure \ref{fig:illus}. We
observe low accuracy in word learning when the perceptual distance
between the labels is small relative to the uncertainty/tolerance with
which these labels are encoded. For example, in Stager and Werker's
original experiment, children are supposed to associate label 1
(``bih'') and label 2 (``dih'') with object 1 and object 2,
respectively. Though children could learn that the label ``bih'' is a
better match to object 1 than ``dih'', they could still judge the sound
``dih'' as a plausible instance of the label ``bih'', thanks to the
relatively large tolerance of the encoding.

Accuracy in word learning increases when the perceived distance between
the labels is large relative to the tolerance of their encoding. This
improvement can occur in the same developmental stage if the perceptual
distance between labels is enhanced through using different-sounding
labels (e.g., ``lif'' vs. ``neem'' instead of ``bih'' and ``dih'') or
through using additional cues (Rost \& McMurray, 2009; Thiessen, 2007).
Besides, accuracy can improve across development through the refinement
of the encoding.

Building on this intuition, the current work proposes a precise
probabilistic model, which we use to both account for previous
experimental findings, and to make new predictions that have not been
tested before. Using new data collected from both preschool children and
adults, we show that the model can explain various patterns of learning
both within the same age and across development.

\begin{CodeChunk}
\begin{figure}[H]

{\centering \includegraphics{figs/illus-1}

}

\caption{\label{fig:illus} An illustration of the probabilistic/continuous account using simulated data. Two words are represented as distributions over the perceptual space. Here accuracy expresses the extent to which a given sound instance indicates the word on the left (red bar), as opposed to the word on the right (green bar). The values vary between 0.5 (total overlap between the distributions) and 1 (no overlap). Accuracy is low when the perceptual distance between labels is small relative to the variance. Accuracy increases when the perceptual distance is enhanced (through disambiguation), or when the variance decreases (e.g., through development).}\label{fig:illus}
\end{figure}
\end{CodeChunk}

\section{Model}\label{model}

\subsection{Task}\label{task}

We model the word learning task introduced by Stager \& Werker (1997),
and a testing method similar to the one used by Yoshida et al. (2009).
In this task, participants are first exposed to the association between
pairs of nonsense words (e.g., ``lif''/``neem'') and pairs of objects.
After this exposure phase, participants perform a series of
two-alternative forced choices. In each testing trial, one of the two
sounds is uttered (e.g., ``lif'') and participants choose the
corresponding object from the two alternatives. An overview of the task
is shown in Figure \ref{fig:task}.

\begin{CodeChunk}
\begin{figure}[t]

{\centering \includegraphics{figs/task-1}

}

\caption{\label{fig:task}An overview of the task used in this study.}\label{fig:task}
\end{figure}
\end{CodeChunk}

\subsection{Probabilistic structure}\label{probabilistic-structure}

Our model consists of a set of variables describing the general process
of spoken word recognition in a referential situation. These variables
are related in a way that reflects the simple generative scenario
represented graphically in Figure \ref{fig:model}. When a speaker utters
a sound in the presence of an object, the observer assumes that the
object \(o\) activated the concept \(C\) in the speaker's mind. The
concept prompted the corresponding label \(L\). Finally, the label was
physically instantiated by the sound \(s\).

A similar probabilistic structure was used by Lewis \& Frank (2013) to
model concept learning, and by Hofer \& Levy (2017) to model spoken word
learning. However, the first study assumed that the sounds are heard
unambiguously, and the second assumed the concepts are observed
unambiguously. In our model, we assume there to be ambiguity at the
level of both the labels and the concepts. For simplicity, we assume
that the probability of membership of objects and sounds to concepts and
labels, respectively, are normally distributed:

\[ p(o | C) \sim  \mathcal{N}(\mu_C, \sigma^2_C) \]
\[ p(s| L) \sim  \mathcal{N}(\mu_L, \sigma^2_L) \] Finally, we assume
there to be one-to-one mappings between concepts and labels, and that
observers have successfully learned these mappings during the exposure
phase: \[
P(L_i|C_j) = 
\begin{cases}
  1 & \text{if  }  i=j \\  
  0  & \text{if  }  i\neq j 
\end{cases}
\]

\subsection{Simulations}\label{simulations}

During the testing phase (see Figure \ref{fig:task}), participants are
presented with one target sound \(s_T \in\{s_1, s_2\}\) and with both
objects \(o_1\) and \(o_2\). In order to make a choice, we need to
determine which object is more probable under the target sound \(s_T\).
In other words, we need to compare the probabilities \(P(o_1|s_T)\) and
\(P(o_2|s_T)\). The values of these probabilities can be computed by
summing over all possible concepts and labels:
\[P(o|s)=\sum_{C,L} P(o, C, L| s) \propto \sum_{C,L} P(o, C, L, s) \]
The joint probability \(P(o, C, L, s)\) is obtained by factoring the
Bayesian network in Figure \ref{fig:model}:
\[P(o,C,L,s) = P(s|L)P(L|C)P(C|o)P(o) \] which could be transformed
using Bayes rule into:

\[P(o,C,L,s) = P(s|L)P(L|C)P(o|C)P(C) \] Finally, assuming that the
concepts' prior probability is uniformly
distributed\footnote{This is a reasonable assumption given the similarity of the concepts used in each naming situation. See the stimuli sub-section in the Experiment below.},
we obtain the following expression, where all conditional dependencies
are now well defined:

\[
P(o|s) = \frac{\sum_{C,L} P(s|L)P(o|C)P(L|C)}{\sum_{o} \sum_{C,L} P(s|L)P(o|C)P(L|C)}
\]

From this general expression we derive the exact analytical formula,
which expresses the probability of accurate responses in the testing
phase of the task:

\begin{equation}
P(o_T|s_T)= \frac{1 + e^{-(\Delta s^2 + \Delta o^2) /2\sigma^2}}{1 + e^{-(\Delta s^2 + \Delta o^2) /2\sigma^2} + e^{-\Delta s^2 /2\sigma^2} + e^{-\Delta o^2 /2\sigma^2 }}
\end{equation}

\begin{CodeChunk}
\begin{figure}[t]

{\centering \includegraphics{figs/model-1}

}

\caption{\label{fig:model}Graphical representation of our model. Circles indicate random variables (shading indicates observed variables). The squares indicates fixed model parameters.}\label{fig:model}
\end{figure}
\end{CodeChunk}

In order to have a more quantitative understanding of the model, we
simulate the values of the predicted accuracy (Expression 1) as a
function of the perceptual distance between the sounds \(\Delta s\). We
used as parameters the two remaining variables, i.e., the visual
distance between the semantic referents \(\Delta o\) and the standard
deviation of the distributions \(p(s| L)\) and \(p(o | C)\) (which, in
this simulation, are assumed to have similar values, i.e.,
\(\sigma =\sigma_C \approx \sigma_L\)). The simulations are shown in
Figure \ref{fig:simulation}.

\begin{CodeChunk}
\begin{figure*}[h]

{\centering \includegraphics{figs/simulation-1}

}

\caption{\label{fig:simulation}The predicted probability of accurate responses in the testing phase, as a function the perceptual distance between the sounds. Colors indicate different values of the standard deviation which we assume is common to both the labels' and concepts' probabilistic representations. Panels represent graphs using different values of the visual distance between the objects.}\label{fig:simulation}
\end{figure*}
\end{CodeChunk}

The simulations explain some previousely documented facts, and make new
predictions:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\item
  For fixed values of \(\Delta o\) and \(\sigma\), the probability of
  accurate responses increases as a function of \(\Delta s\). This
  pattern accounts for the fact that similar sounds are generally more
  challenging to learn than different sounds for both children (Stager
  \& Werker, 1997) and adults (Pajak et al., 2016).
\item
  For fixed values of \(\Delta s\) and \(\Delta o\), accuracy increases
  when the representational uncertainty (characterized with \(\sigma\))
  decreases. This fact may explain development, i.e., younger children
  have noisier representations (see Swingley, 2007; Yoshida et al.,
  2009), which leads to lower word recognition accuracy, especially for
  similar-sounding words.
\item
  For fixed values of \(\Delta s\) and \(\sigma\), accuracy increases
  with the visual distance between the semantic referents \(\Delta o\).
  This is a new prediction that our model makes. Previous work studied
  the effect of several bottom-up and top-down properties in
  disambiguating similar sounding words (Dautriche, Swingley, \&
  Christophe, 2015; Fennell \& Waxman, 2010; Rost \& McMurray, 2009;
  Thiessen, 2007), but no previous study tested the effect of the visual
  distance between the semantic referents.
\end{enumerate}

To sum, we introduced a model that accounts for some qualitative
learning patterns observed in previous studies, and makes a new
prediction. In the experiment below, we test whether the model makes
accurate \emph{quantitative} predictions by fitting it to new
experimental data collected from preschool children and adults.

\section{Experiment}\label{experiment}

In this experiment, we tested participants in the word learning task
introduced above (Figure \ref{fig:task}). We explored all three
parameters of the model. In particular, both the sound similarity
(\(\Delta s\)) and object similarity (\(\Delta o\)) were varied
simultaneously in a within-subject design. Besides, two age groups
(preschool children and adults) were tested on the same task to explore
whether development can be characterized with the uncertainty parameter,
\(\sigma\), in the probabilsitic representations.

\subsection{Methods}\label{methods}

\subsubsection{Participants}\label{participants}

We planned to recruit a sample of 60 children ages 4-5 years from the
Bing Nursery School on Stanford University's campus. So far, we
collected data from N=47 children (mean age= months, F=). An additional
28 children participated but were removed from analyses because they
were not above chance on the catch trials (as was specified in the
pre-registration\footnote{https://osf.io/jrh38/}). We also collected a
sample of N=20 adults on Amazon Mechanical Turk (this is a pilot, I will
plan for 60 participants). N=2 adult participants were excluded because
of low scores on the catch trials (see pre-registration). (I still need
to write the code that derives these numbers automatically from the
data).

\subsubsection{Stimuli and similarity
rating}\label{stimuli-and-similarity-rating}

The sound stimuli were generated using the MBROLA Speech Synthesizer
(Dutoit, Pagel, Pierret, Bataille, \& Van der Vrecken, 1996). We
generated three kinds of nonsense word pairs which varied in their
degree of similarity to English speakers: 1) ``different'':
``lif''/``neem'' and ``zem''/``doof'', 2) ``intermediate'':
``aka''/``ama'' and ``ada''/``aba'', and 3) ``similar'' non-English
minimal pairs: ``ada''/``a\textipa{d\super h}a'' (in hindi) and
``a\textipa{Q}a''/``a\textipa{\textcrh}a'' (in arabic).

As for the objects, we used the Dynamic Stimuli javascript
library\footnote{https://github.com/erindb/stimuli} which allowed us to
generate objects in four different categories: ``tree'', ``bird'',
``bug'', and ``fish''. These categories are supposed to be naturally
occurring kinds that might be seen on an alien planet. In each category,
we generated ``different'', ``intermediate'' and ``similar'' pairs by
manipulating a continuous property controlling features of the
category's shape (e.g, body stretch or head fatness).

In a separate survey, \(N=20\) participants recruited on Amazon
Mechanical Turk evaluated the similarity of each sound and object pair
on a 7-point scale. We processed the resulting data so that it can be
used in the model. We normalized the values with respect to the most
distant level, and we scaled them within the range {[}0,1{]}. Processed
data are shown in Figure \ref{fig:stim}, for each stimuli group. This
data will be used in the models as the perceptual distance of sound
pairs (\(\Delta s\)) and object pairs (\(\Delta o\)).

\begin{CodeChunk}
\begin{figure}[h]

{\centering \includegraphics{figs/stim-1}

}

\caption{\label{fig:stim}Normalized distances for both sound and object pairs used in this study.}\label{fig:stim}
\end{figure}
\end{CodeChunk}

\subsubsection{Design}\label{design}

Each age group saw only two of the three levels of similarity described
in the previous sub-section: ``different'' vs. ``intermediate'' for
preschoolers, and ``intermediate'' vs. ``similar'' for adults. We made
this choice because, on the one hand, adults were at ceiling with
``different'' sounds/objects, and on the other hand, children were at
chance with the ``similar'' sounds/objects. That said, this difference
in the level of similarity is accounted for in the model through using
the appropriate perceptual distance used in each age group (Figure
\ref{fig:stim}).

The experiment consisted of four conditions which involved, each, one
pair of sounds-objects associations. These conditions were constructed
by crossing the sound's degree of similarity with the object's degree of
similarity leading to a 2x2 factorial design in each age group. Besides
the 4 conditions, we also tested participants on a fifth catch condition
which was similar in its structure to the other ones, but was used only
to select participants who were able to follow the instructions and show
minimal learning.

\begin{CodeChunk}
\begin{figure*}[h]

{\centering \includegraphics{figs/all_data-1}

}

\caption{\label{fig:data_all}Accuracy of novel word recognition as as a function of the sound distance, the object distance, and the age group (preschool children vs. Adults). Experimental results are shown on the left. Predictions from Model 1 (one free parameter) and Model 2 (two free parameters) are shown in the middle and on the right, respectively.}\label{fig:all_data}
\end{figure*}
\end{CodeChunk}

\subsubsection{Procedure}\label{procedure}

Preschoolers were asked if they would be willing to play a game on a
tablet with the experimenter and were informed that they could stop
playing at any time. The experimenter explained that the game consisted
in learning some words spoken in an alien planet. The experiment began
with two simple examples (not included in the analysis), and in these
examples children were given feedback from the experimenter so as to
make sure they correctly understood the structure of the task. After the
introduction and examples, children were tested in a sequence of five
conditions: the four experimental conditions plus the catch condition.
In each condition, participants saw a first block of four exposure
trials followed by four testing trials, and a second block of two
exposure trials (for memory refreshment) followed by an additional four
testing trials.

In the exposure trials, children saw two objects associated with their
corresponding sounds. We presented the first object on the left side of
the tablet's screen simultaneously with the corresponding sound. The
second sound-object association followed on the other side of the screen
after 500ms. For both objects, visual stimuli were present for the
duration of the sound clip (∼800ms). In the testing trials, children saw
both objects simultaneously and heard only one sound. They completed the
trial by selecting which of the two objects corresponded to the sound.
They responded by touching one of the pictures on the tablet.

The object-sound pairings were randomized across participants, as was
the order of the conditions (except for the catch condition which was
always placed in the middle of the testing sequence). We also randomized
the on-screen position (left vs.~right) of the two pictures on each
testing trial.

The procedure in the adult experiment was identical to that of preschool
children, except that preschoolers were accompanied by an experimenter
and used a tablet, whereas adults used their local computers to complete
the experiment online.

\subsubsection{Model fitting}\label{model-fitting}

We fit the probability function (equation 2) to the participants'
responses in each age group. The values of \(\Delta s\) and \(\Delta o\)
were set based on data from the similarity judgment task (Figure
\ref{fig:stim}). We used two models: \textbf{model 1} fit only one
parameter (\(\sigma = \sigma_C =\sigma_L\)), and \textbf{model 2} fit
two parameters (\(\sigma_C \neq \sigma_L\)). The values of the
parameters were derived using weighted least-squares estimates.

\subsection{Results}\label{results}

We first analyzed the experimental results shown in Figure
(\ref{fig:data_all}, left), using a mixed-effects logistic regression
with sound and object distances as fixed effects, and with a maximal
random effects structure (Barr, Levy, Scheepers, \& Tily, 2013). Results
are shown in Table 1. We found a main effect of sound distance on the
accuracy of learning in both children and adults, thus replicating
previous findings. We also found a main effect of object distance, thus
confirming the new prediction of our model.

\begin{table}[!htbp] \centering 
  \caption{Estimates of predictor coefficients (and their standard errors) by age group in the regression model} 
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}}lcc} 
\\[-1.8ex]\hline \\[-1.8ex] 
 & Children & Adults \\ 
\hline \\[-1.8ex] 
 (Intercept) & 0.426$^{*}$ (0.199) & 3.114$^{**}$ (1.015) \\ 
  Sound  & 0.272$^{**}$ (0.100) & 2.320$^{*}$ (0.981) \\ 
  Object & 0.315$^{*}$ (0.137) & 2.133$^{*}$ (0.952) \\ 
  Sound x Object & 0.151 (0.097) & 1.821 (0.976) \\ 
 \hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.05; $^{**}$p$<$0.01; $^{***}$p$<$0.001} \\ 
\end{tabular} 
\end{table}

Figures (\ref{fig:data_all}, middle and right graphs) show the
predictions of the models. Both model 1 and model 2 fit reasonably well
the experimental data in both children and adults. In particular, they
both correctly predict the relative recognition accuracy across
conditions: the pair of words that differ on both the object and sound
levels were the easiest to learn, followed by the pairs of words that
differ on only one level, then the pair of words that are similar on
both levels. Note, however, that both models predict a slight
interaction between sounds and objects, but such an interaction was not
significant in the the experimental data (Table 1).

For Model 1, the fitted uncertainty parameter was \(\sigma =\) 0.63
{[}0.53, 0.73{]}\footnote{95 \% Confidence intervals.} for preschoolers,
and \(\sigma =\) 0.16 {[}0.12, 0.19{]} for adults. As predicted, the
uncertainty parameter was larger for children than it was for adults.
The model explained the majority of the variance (\(R^2=\) 0.94). For
model 2, children had a label-specific uncertainty of \(\sigma_S =\) 0.9
{[}0.68, 1.11{]}, and a concept-specific uncertainty of \(\sigma_C =\)
0.29 {[}0.1, 0.49{]}. Adults had a label-specific noise of
\(\sigma_S =\) 0.16 {[}0.05, 0.28{]}, and a concept-specific noise of
\(\sigma_C =\) 0.14 {[}0.05, 0.23{]}. The model explained almost all the
variance (\(R^2=\) 0.96). Interestingly, despite the fact that Model 1
used only one degree of freedom, it captures the data variance almost as
well as Model 2 (which used two degrees of freedom).

\section{General Discussion}\label{general-discussion}

Since the seminal work of Stager \& Werker (1997) showing a discrepancy
between children's speech perception abilities and their word learning
skills, researchers have debated about the source of such a discrepancy.
In this work, we were more interested in the development of this
phenomenon, regardless of its exact origin. Building on some previous
findings and discussions (Swingley, 2007; Yoshida et al., 2009), we
proposed a model which assumes that subtle perceptual features are
correctly encoded during word meaning learning, but that the precision
of this encoding vary across development. We compared the predictions to
new data which we collected from preschool children and adults. We found
that the probabilistic model captured not only the patterns of
development, but also various learning patterns within the same age
group.

The model made a new prediction which we tested experimentally for the
first time: learning similar words is not only modulated by to
similarity of their phonological forms, but also by the visual
similarity of their semantic referents. More generally, since visual
similarity is an early organizing feature in the semantic domain (e.g.,
Wojcik \& Saffran, 2013), our finding suggests that children may
prioritize the acquisition of words that are quite distant in the
semantic space. This suggestion is supported by recent findings based on
the investigation of early vocabulary growth (Engelthaler \& Hills,
2017; Sizemore, Karuza, Giusti, \& Bassett, 2018). That said, further
work is needed to explore the effect on word learning of other semantic
dimensions that could be encoded by children (e.g.,
conceptual/functional features).

To conclude, this paper proposes a computational model for the
probabilistic/continuous account, and the findings do endorse this
account as a valid framework for the study of similar word learning and
the mechanism of its development. More generally, together with previous
modeling efforts (e.g., Apfelbaum \& McMurray, 2011; Hofer \& Levy,
2017), our work highlight the crucial role computation models can play
in theory building through both a compact organization of existing
knowledge and through generating new principled hypothesis.

\vspace{1em}

\fbox{\parbox[b][][c]{7.3cm}{\centering All data and code for these analyses are available at\ \url{https://github.com/afourtassi/networks}}}
\vspace{1em}

\section{Acknowledgements}\label{acknowledgements}

This work was supported by a post-doctoral grant from the Fyssen
Foundation, NSF \#1528526, and NSF \#1659585.

\section{References}\label{references}

\setlength{\parindent}{-0.1in} \setlength{\leftskip}{0.125in} \noindent

\hypertarget{refs}{}
\hypertarget{ref-apfelbaum2011}{}
Apfelbaum, K., \& McMurray, B. (2011). Using variability to guide
dimensional weighting: Associative mechanisms in early word learning.
\emph{Cognitive Science}, \emph{35}(6).

\hypertarget{ref-barr2013}{}
Barr, D., Levy, R., Scheepers, C., \& Tily, H. (2013). Random effects
structure for confirmatory hypothesis testing: Keep it maximal.
\emph{Journal of Memory and Language}, \emph{68}(3).

\hypertarget{ref-dautriche2015}{}
Dautriche, I., Swingley, D., \& Christophe, A. (2015). Learning novel
phonological neighbors: Syntactic category matters. \emph{Cognition},
\emph{143}.

\hypertarget{ref-dutoit1996}{}
Dutoit, T., Pagel, V., Pierret, N., Bataille, F., \& Van der Vrecken, O.
(1996). The mbrola project: Towards a set of high quality speech
synthesizers free of use for non commercial purposes. In
\emph{Proceedings of ICSLP} (Vol. 3). IEEE.

\hypertarget{ref-engelthaler2017}{}
Engelthaler, T., \& Hills, T. T. (2017). Feature biases in early word
learning: Network distinctiveness predicts age of acquisition.
\emph{Cognitive Science}, \emph{41}.

\hypertarget{ref-fennell2010}{}
Fennell, C., \& Waxman, S. (2010). What paradox? Referential cues allow
for infant use of phonetic detail in word learning. \emph{Child
Development}, \emph{81}.

\hypertarget{ref-hofer2017}{}
Hofer, M., \& Levy, R. (2017). Modeling Sources of Uncertainty in Spoken
Word Learning. In \emph{Proceedings of the 39th Annual Meeting of the
Cognitive Science Society}.

\hypertarget{ref-lewis2013}{}
Lewis, M., \& Frank, M. (2013). An integrated model of concept learning
and word-concept mapping. In \emph{Proceedings of the annual meeting of
the cognitive science society} (Vol. 35).

\hypertarget{ref-pajak2016}{}
Pajak, B., Creel, S., \& Levy, R. (2016). Difficulty in learning
similar-sounding words: A developmental stage or a general property of
learning? \emph{Journal of Experimental Psychology: Learning, Memory,
and Cognition}, \emph{42}(9).

\hypertarget{ref-rost2009}{}
Rost, G., \& McMurray, B. (2009). Speaker variability augments
phonological processing in early word learning. \emph{Developmental
Science}, \emph{12}.

\hypertarget{ref-sizemore2018}{}
Sizemore, A. E., Karuza, E. A., Giusti, C., \& Bassett, D. S. (2018).
Knowledge gaps in the early growth of semantic feature networks.
\emph{Nature Human Behaviour}, \emph{2}(9).

\hypertarget{ref-stager1997}{}
Stager, C., \& Werker, J. (1997). Infants listen for more phonetic
detail in speech perception than in word-learning tasks. \emph{Nature},
\emph{388}(6640).

\hypertarget{ref-swingley2007}{}
Swingley, D. (2007). Lexical exposure and word-form encoding in
1.5-year-olds. \emph{Developmental Psychology}, \emph{43}(2).

\hypertarget{ref-thiessen2007}{}
Thiessen, E. (2007). The effect of distributional information on
children's use of phonemic contrasts. \emph{Journal of Memory and
Language}, \emph{56}.

\hypertarget{ref-werker1984}{}
Werker, J., \& Tees, R. (1984). Cross-language speech perception:
Evidence for perceptual reorganization during the first year of life.
\emph{Infant Behavior and Development}, \emph{7}.

\hypertarget{ref-werker2002}{}
Werker, J., Fennell, C., Corcoran, K., \& Stager, C. (2002). Infants'
ability to learn phonetically similar words: Effects of age and
vocabulary size. \emph{Infancy}, \emph{3}.

\hypertarget{ref-white2013}{}
White, K., Yee, E., Blumstein, S., \& Morgan, J. (2013). Adults show
less sensitivity to phonetic detail in unfamiliar words, too.
\emph{Journal of Memory and Language}, \emph{68}(4).

\hypertarget{ref-wojcik2013}{}
Wojcik, E., \& Saffran, J. (2013). The ontogeny of lexical networks:
Toddlers encode the relationships among referents when learning novel
words. \emph{Psychological Science}, \emph{24}(10).

\hypertarget{ref-yoshida2009}{}
Yoshida, K., Fennell, C., Swingley, D., \& Werker, J. (2009).
14-month-olds learn similar-sounding words. \emph{Developmental
Science}, \emph{12}.

\end{document}