### 2019-01-29 cogsci/cogsci.tex

% Template for Cogsci submission with R Markdown

% Stuff changed from original Markdown PLOS Template
\documentclass[10pt, letterpaper]{article}

\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{float}
\usepackage{caption}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}

% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% hyperref package, useful for hyperlinks
\usepackage{hyperref}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% Sweave(-like)
\usepackage{fancyvrb}
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
\newenvironment{Schunk}{}{}
\DefineVerbatimEnvironment{Code}{Verbatim}{}
\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
\newenvironment{CodeChunk}{}{}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace}
%\doublespacing

% % Text layout
% \topmargin 0.0cm
% \oddsidemargin 0.5cm
% \evensidemargin 0.5cm
% \textwidth 16cm
% \textheight 21cm

\title{The role of continuous developmental change in explaining
discontinuities in word learning}

\usepackage{tipa}
\usepackage[rightcaption]{sidecap}

\author{{\large \bf Abdellah Fourtassi} \\ \texttt{afourtas@stanford.edu} \\ Department of Psychology \\ Stanford University \And {\large \bf Sophie Regan} \\ \texttt{sregan20@stanford.edu} \\ Department of Psychology \\ Stanford University \And {\large \bf Michael C. Frank} \\ \texttt{mcfrank@stanford.edu} \\ Department of Psychology \\ Stanford University}

\begin{document}

\maketitle

\begin{abstract}
Cognitive development is often characterized in term of discontinuities.
Nevertheless, in some cases, this characterization can be problematic
and unnecessary. In this paper, we show that a phenomenon that appears
stage-like, can be well described by a continuous account. We used as a
case study the finding by Stager and Werker (1997) that children's early
ability to distinguish similar sounds does not automatically translate
into word learning skills. Early explanations proposed that children may
not be able to encode subtle phonetic contrasts when learning novel word
meanings, thus suggesting a discontinuous/stage-like pattern of
development. However, later work has revealed (e.g., through using
simpler testing methods) that children do encode such contrasts, thus
favoring a continuous pattern of development. Here we propose a precise
probabilistic model describing how development may proceed in a
continuous fashion across the lifespan. The model accounts for
previously documented facts and provides new predictions. We collected
data from preschool children and adults, and we showed that the model
can explain various patterns of learning both within the same age and
across development. The findings suggest that other aspects of cognitive
development which are typically thought of as discontinuities, may
emerge from simpler, continuous mechanisms.

\textbf{Keywords:}
word learning, cognitive development, computational modeling
\end{abstract}

\section{Introduction}\label{introduction}

Cognitive development is sometimes characterized in terms of a
succession of discontinuous stages (Piaget, 1954). Such characterization
may be problematic as it often renders the study of the learning
mechanism quite challenging. Indeed, it is unclear how a precise
algorithm----whose steps are gradual and logically related----can
account for transitions between stages involving qualitatively different
representations (e.g., Carey, 2009).

Nevertheless, at least in some cases, development may only appear to be
stage-like. This appearance can be due, for example, to the use of a
cognitively-demanding task which may mask learning, or to the use of
hard statistical thresholding (in particular, p-value \textless{} 0.05)
which creates a spurious dichotomy between success and failure in
observing a given behavior. In such cases, positing discontinuous stages
is unnecessary. Instead, as we shall see in this paper, a continuous
model----involving similar representations across the lifespan----may
provide a simpler and more transparent account of development.

We use a case study from word learning literature. Stager \& Werker
(1997) first showed that children's early ability to distinguish similar
sounds (``dih'' vs. ``bih'') does not automatically translate into word
learning skills (i.e., associating these sounds with different
meanings). Subsequent research has focused on proposing possible
explanations for this observed gap between speech perception and word
learning (e.g., Fennell \& Waxman, 2010; Hofer \& Levy, 2017; Rost \&
McMurray, 2009; Stager \& Werker, 1997).

In this work we focus, rather, on the development of this phenomenon. In
fact, regardless of the exact explanation, it is generally accepted that
by around 17 m.o, children succeed in the same task (Werker, Fennell,
Corcoran, \& Stager, 2002). How does development proceed? Early accounts
assumed that children encode words in a binary way: they either fail or
succeed in encoding the relevant phonetic details (simultaneously with
the meanings). This account suggested a discontinuous/stage-like pattern
of development whereby younger children fail to encode the contrastive
phonetic detail, whereas older children succeed.

Nevertheless, subsequent findings have suggested otherwise. On the one
hand, 14-month-olds--who typically fail in the original task--succeed
when an easier testing method is used, even under the same learning
conditions (Yoshida, Fennell, Swingley, \& Werker, 2009). They also
succeed when uncertainty is mitigated via disambiguating cues (e.g.,
Thiessen, 2007). On the other hand, adults show patterns of learning
similar to those shown by 14-month-olds when the task is more
challenging and when similarity between words increases (Pajak, Creel,
\& Levy, 2016; White, Yee, Blumstein, \& Morgan, 2013).

These evidence point towards another scenario, where the representations
are encoded in a probabilistic (rather than binary) way, and where
development is continuous, rather than stage-like (Pajak et al., 2016;
Swingley, 2007; Yoshida et al., 2009). On this account, correct
representations are learned early in development, but these
representations are encoded with higher uncertainty in younger children,
leading to apparent failure in relatively demanding tasks. Development
is a continuous process whereby the initial noisy representations become
more precise. Crucially, more precise representation are still
imperfect: Even adults show low accuracy learning when the sounds are
subtle enough, e.g., non-native sounds (Pajak et al., 2016).

We provide an intuitive illustration of how such an account explains
patterns of learning and development in Figure \ref{fig:illus}. We
observe low accuracy in word learning when the perceptual distance
between the labels is small relative to the uncertainty with which these
labels are encoded. For example, in Stager and Werker's original
experiment, children are supposed to associate label 1 (``bih'') and
label 2 (``dih'') with object 1 and object 2, respectively. Though
children could learn that the label ``bih'' is a better match to object
1 than ``dih'', they could still judge the sound ``dih'' as a plausible
instance of the label ``bih'', thanks to the relatively large
uncertainty of the encoding.

According to this account, accuracy in word learning improves if we
increase either the perceptual distinctiveness of the stimuli (e.g.,
through using different-sounding labels), or the precision of the
encoding itself (e.g., across development). Building on this intuition,
the current work proposes a precise probabilistic model, which we use to
both account for previous experimental findings, and to make new
predictions that have not been tested before. Using new data collected
from both preschool children and adults, we show that the model can
explain various patterns of learning both within the same age and across
development.

\begin{CodeChunk}
\begin{figure}[H]

{\centering \includegraphics{figs/illus-1}

}

\caption{\label{fig:illus} An illustration of the probabilistic/continuous account using simulated data. A word is represented with a distribution over the perceptual space (indicated in red or green). When the uncertainty of the representation is large relative to the distance between the stimuli (top panel), an instance of the red category (indicated with a star) could also be a plausible instance of the green category, hence the low recognition accuracy score. The accuracy increases when the stimuli are less similar (left panel), or when the representation are more precise (right panel).}\label{fig:illus}
\end{figure}
\end{CodeChunk}

\section{Model}\label{model}

\subsection{Probabilistic structure}\label{probabilistic-structure}

Our model consists of a set of variables describing the general process
of spoken word recognition in a referential situation. These variables
are related in a way that reflects the simple generative scenario
represented graphically in Figure \ref{fig:model}. When a speaker utters
a sound in the presence of an object, the observer assumes that the
object \(o\) activated the concept \(C\) in the speaker's mind. The
concept prompted the corresponding label \(L\). Finally, the label was
physically instantiated by the sound \(s\).

\begin{SCfigure}
\centering
\includegraphics[width=1.5in]{figs/model.png}
\caption{Graphical representation of our model. Circles indicate random variables (shading indicates observed variables). The squares indicates fixed model parameters.}
\label{fig:model}
\end{SCfigure}

A similar probabilistic structure was used by Lewis \& Frank (2013) to
model concept learning, and by Hofer \& Levy (2017) to model spoken word
learning. However, the first study assumed that the sounds are heard
unambiguously, and the second assumed the concepts are observed
unambiguously. In our model, we assume there to be ambiguity at the
level of both the labels and the concepts. For simplicity, we assume
that the probability of membership of objects and sounds to concepts and
labels, respectively, are normally distributed:

\[ p(o | C) \sim  \mathcal{N}(\mu_C, \sigma^2_C) \]
\[ p(s| L) \sim  \mathcal{N}(\mu_L, \sigma^2_L) \] Finally, we assume
there to be one-to-one mappings between concepts and labels, and that
observers have successfully learned these mappings during the exposure
phase: \[
P(L_i|C_j) = 
\begin{cases}
  1 & \text{if  }  i=j \\  
  0  & \text{if  }  i\neq j 
\end{cases}
\]

\subsection{Inference}\label{inference}

The learner hears a sound \(s\) and has to decided which object \(o\)
provides an optimal match to this sound. To this end, the learner has to
compute the probability \(P(o|s)\) for all possible objects. This
probability can be computed by summing over all possible concepts and
labels:
\[P(o|s)=\sum_{C,L} P(o, C, L| s) \propto \sum_{C,L} P(o, C, L, s) \]
The joint probability \(P(o, C, L, s_T)\) is obtained by factoring the
Bayesian network in Figure \ref{fig:model}:
\[P(o,C,L,s) = P(s|L)P(L|C)P(C|o)P(o) \] which could be transformed
using Bayes rule into:

\[P(o,C,L,s) = P(s|L)P(L|C)P(o|C)P(C) \] Finally, assuming that the
concepts' prior probability is uniformly
distributed\footnote{This is a reasonable assumption given the similarity of the concepts used in each naming situation. See the stimuli sub-section in the Experiment below.},
we obtain the following expression, where all conditional dependencies
are now well defined:

\begin{equation}
P(o|s) = \frac{\sum_{C,L} P(s|L)P(o|C)P(L|C)}{\sum_{o} \sum_{C,L} P(s|L)P(o|C)P(L|C)}
\end{equation}

\subsection{Task and model
predictions}\label{task-and-model-predictions}

\begin{CodeChunk}
\begin{figure}[t]

{\centering \includegraphics{figs/task-1}

}

\caption{\label{fig:task}An overview of the task used in this study.}\label{fig:task}
\end{figure}
\end{CodeChunk}

We use the model to predict performance in the word learning task
introduced by Stager \& Werker (1997), and a testing method similar to
the one used by Yoshida et al. (2009). In this task, participants are
first exposed to the association between pairs of nonsense words (e.g.,
``lif''/``neem'') and pairs of objects. After this exposure phase,
participants perform a series of test trials. In each of these trials,
one of the two sounds is uttered (e.g., ``lif'') and participants choose
the corresponding object from the two alternatives. An overview of the
task is shown in Figure \ref{fig:task}.

We use the general Equation 1 to derive the exact analytical expression
for the probability of accurate responses \(p(o_T | s)\) (target object
\(o_T\) given a sound \(s\)) in the simple case of two-alternative
forced choice in the testing phase of our experimental task:

\begin{equation}
P(o_T|s)= \frac{1 + e^{-(\Delta s^2/2\sigma_L^2+ \Delta o^2/2\sigma_C^2)}}{1 + e^{-(\Delta s^2/2\sigma_L^2+ \Delta o^2/2\sigma_C^2)}+ e^{-\Delta s^2 /2\sigma_L^2} + e^{-\Delta o^2 /2\sigma_C^2 }}
\end{equation}

Figure \ref{fig:simulation} show simulations of the predicted accuracy
(Expression 2) as a function of the distinctiveness parameters
(\(\Delta s\) and \(\Delta o\)) and the precision parameter, i.e., the
variance of the distributions \(p(s| L)\) and \(p(o | C)\) (which, for
ease of the simulations' interpretation, are assumed to have similar
values, i.e., \(\sigma =\sigma_C \approx \sigma_L\)).

\begin{CodeChunk}
\begin{figure*}[h]

{\centering \includegraphics{figs/simulation-1}

}

\caption{\label{fig:simulation}The predicted probability of accurate responses in the testing phase as a function of stimuli distinctiveness $\Delta s$ (x-axis) and $\Delta o$ (different panels) and representation precision $\sigma$ (color). Dashed line represents chance.}\label{fig:simulation}
\end{figure*}
\end{CodeChunk}

The simulations explain some previously documented facts, and make new
predictions:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\item
  For fixed values of \(\Delta o\) and \(\sigma\), the probability of
  accurate responses increases as a function of \(\Delta s\). This
  pattern accounts for the fact that similar sounds are generally more
  challenging to learn than different sounds for both children (Stager
  \& Werker, 1997) and adults (Pajak et al., 2016).
\item
  For fixed values of \(\Delta s\) and \(\Delta o\), accuracy increases
  when the representational uncertainty (characterized with \(\sigma\))
  decreases. This fact may explain development, i.e., younger children
  have noisier representations (see Swingley, 2007; Yoshida et al.,
  2009), which leads to lower word recognition accuracy, especially for
  similar-sounding words.
\item
  For fixed values of \(\Delta s\) and \(\sigma\), accuracy increases
  with the visual distance between the semantic referents \(\Delta o\).
  This is a new prediction that our model makes. Previous work studied
  the effect of several bottom-up and top-down properties in
  disambiguating similar sounding words (e.g., Fennell \& Waxman, 2010;
  Rost \& McMurray, 2009; Thiessen, 2007), but no previous study tested
  the effect of the visual distance between the semantic referents.
\end{enumerate}

In sum, we introduced a model that accounts for some qualitative
learning patterns observed in previous studies, and makes a new
prediction. In the experiment below, we test whether the model makes
accurate quantitative predictions by fitting it to new experimental data
collected from preschool children and adults.

\section{Experiment}\label{experiment}

In this experiment, we tested participants in the word learning task
introduced above (Figure \ref{fig:task}). We explored all three
parameters of the model. In particular, both the sound similarity
(\(\Delta s\)) and object similarity (\(\Delta o\)) were varied
simultaneously in a within-subject design. Two age groups (preschool
children and adults) were tested on the same task to explore whether
development can be characterized with the uncertainty parameter,
\(\sigma\), in the probabilistic representations. The Experiment and the
predictions were pre-registered.

\subsection{Methods}\label{methods}

\subsubsection{Participants}\label{participants}

We planned to recruit a sample of 60 children ages 4-5 years from the
Bing Nursery School on Stanford University's campus. So far, we
collected data from N=47 children (mean age= months, F=). An additional
28 children participated but were removed from analyses because they
were not above chance on the catch trials (as was specified in the
pre-registration\footnote{https://osf.io/jrh38/}). We also collected a
sample of N=20 adults on Amazon Mechanical Turk (this is a pilot, I will
plan for 60 participants). N=2 adult participants were excluded because
of low scores on the catch trials (see pre-registration). (I still need
to write the code that derives these numbers automatically from the
data).

\subsubsection{Stimuli and similarity
rating}\label{stimuli-and-similarity-rating}

The sound stimuli were generated using the MBROLA Speech Synthesizer
(Dutoit, Pagel, Pierret, Bataille, \& Van der Vrecken, 1996). We
generated three kinds of nonsense word pairs which varied in their
degree of similarity to English speakers: 1) ``different'':
``lif''/``neem'' and ``zem''/``doof'', 2) ``intermediate'':
``aka''/``ama'' and ``ada''/``aba'', and 3) ``similar'' non-English
minimal pairs: ``ada''/``a\textipa{d\super h}a'' (in hindi) and
``a\textipa{Q}a''/``a\textipa{\textcrh}a'' (in arabic).

As for the objects, we used the Dynamic Stimuli javascript
library\footnote{https://github.com/erindb/stimuli} which allowed us to
generate objects in four different categories: ``tree'', ``bird'',
``bug'', and ``fish''. These categories are supposed to be naturally
occurring kinds that might be seen on an alien planet. In each category,
we generated ``different'', ``intermediate'' and ``similar'' pairs by
manipulating a continuous property controlling features of the
category's shape (e.g, body stretch or head fatness).

In a separate survey, \(N=20\) participants recruited on Amazon
Mechanical Turk evaluated the similarity of each sound and object pair
on a 7-point scale. We processed the resulting data so that it can be
used in the model. We normalized the values with respect to the most
distant level, and we scaled them within the range {[}0,1{]}. Processed
data are shown in Figure \ref{fig:stim}, for each stimuli group. This
data will be used in the models as the perceptual distance of sound
pairs (\(\Delta s\)) and object pairs (\(\Delta o\)).

\begin{CodeChunk}
\begin{figure}[h]

{\centering \includegraphics{figs/stim-1}

}

\caption{\label{fig:stim}Distances for both sound and object pairs from an adult norming study. Error bars represent 95\% confidence intervals.}\label{fig:stim}
\end{figure}
\end{CodeChunk}

\subsubsection{Design}\label{design}

Each age group saw only two of the three levels of similarity described
in the previous sub-section: ``different'' vs. ``intermediate'' for
preschoolers, and ``intermediate'' vs. ``similar'' for adults. We made
this choice because, on the one hand, adults were at ceiling with
``different'' sounds/objects, and on the other hand, children were at
chance with the ``similar'' sounds/objects. That said, this difference
in the level of similarity is accounted for in the model through using
the appropriate perceptual distance used in each age group (Figure
\ref{fig:stim}).

The experiment consisted of four conditions which involved, each, one
pair of sounds-objects associations. These conditions were constructed
by crossing the sound's degree of similarity with the object's degree of
similarity leading to a 2x2 factorial design in each age group. Besides
the 4 conditions, we also tested participants on a fifth catch condition
which was similar in its structure to the other ones, but was used only
to select participants who were able to follow the instructions and show
minimal learning.

\begin{CodeChunk}
\begin{figure*}[h]

{\centering \includegraphics{figs/all_data-1}

}

\caption{\label{fig:data_all}Accuracy of novel word recognition as as a function of the sound distance, the object distance, and the age group (preschool children vs. Adults). We show both experimental results (top) and model predictions (bottom). Error bars represent 95\% confidence intervals.}\label{fig:all_data}
\end{figure*}
\end{CodeChunk}

\subsubsection{Procedure}\label{procedure}

Preschoolers were tested at the nursery school using a tablet, whereas
adults used their local computers to complete the same experiment
online. Participants were tested in a sequence of five conditions: the
four experimental conditions plus the catch condition. In each
condition, participants saw a first block of four exposure trials
followed by four testing trials, and a second block of two exposure
trials (for memory refreshment) followed by an additional four testing
trials.

In the exposure trials, participants saw two objects associated with
their corresponding sounds. We presented the first object on the left
side of the tablet's screen simultaneously with the corresponding sound.
The second sound-object association followed on the other side of the
screen after 500ms. For both objects, visual stimuli were present for
the duration of the sound clip (∼800ms). In the testing trials,
participants saw both objects simultaneously and heard only one sound.
They completed the trial by selecting which of the two objects
corresponded to the sound. The object-sound pairings were randomized
across participants, as was the order of the conditions (except for the
catch condition which was always placed in the middle of the testing
sequence). We also randomized the on-screen position (left vs.~right) of
the two pictures on each testing trial.

\subsection{Results}\label{results}

We first analyzed the experimental results shown in Figure
(\ref{fig:data_all}, top), using a mixed-effects logistic regression
with sound distance, object distance and age group as fixed effects, and
with a maximal random effects structure (allowing us to take into
account the full nested structure of our data) (Barr, Levy, Scheepers,
\& Tily, 2013). We found main effects for all the fixed effects, i.e.,
the sound distance (\(\beta =\) 0.68, \(p\) \textless{}
0.001)---replicating previous findings, the object distance (\(\beta =\)
0.6, \(p\) \textless{} 0.001)---confirming the new prediction of our
model, and age group (\(\beta =\) 0.59, \(p\) \textless{}
0.001)---showing that performance improves with age. Besides, we found
two-way interactions between sound and object distances (\(\beta =\)
0.36, \(p\) \textless{}0.05) and between sound distance and age
(\(\beta =\) 0.37, \(p\) \textless{}0.01).

Second, we fit the probability function (Equation 2) to the
participants' responses in each age group. The values of \(\Delta s\)
and \(\Delta o\) were set based on data from the similarity judgment
task (Figure \ref{fig:stim}). Thus, the model has two degrees of freedom
in each group, i.e., \(\sigma_C\) and \(\sigma_L\). Figure
\ref{fig:data_all} (bottom) show the predictions of the model. The model
correctly predicts the relative recognition accuracy across conditions:
the pair of words that differ on both the object and sound levels were
the easiest to learn, followed by the pairs of words that differ on only
one level, then the pair of words that are similar on both levels. At
the quantitative level, the models explained almost all the variance in
the participants' mean responses (\(R^2=\) 0.96).

We also fit a simplified model with only one degree of freedom
(\(\sigma = \sigma_C =\sigma_L\)). This second model is not intended to
test any theoretical hypothesis. We use it, rather, as a baseline to
understand how much of the explanatory power of the model is due to the
extra degree of freedom. This baseline model still captured the
qualitative pattern (graph not shown), and explained the majority of the
variance (\(R^2=\) 0.94). Besides, model comparison using Akaike
information criterion yielded \(AIC(model) - AIC(baseline) =\) -3.79 for
children data and \(AIC(model) - AIC(baseline) =\) 1.62 for adult data.
These numbers suggest that the model remains highly predictive even with
one degree of freedom.

As for the values of the parameters, children had a label-specific
uncertainty of \(\sigma_S =\) 0.9 {[}0.68,
1.11{]}\footnote{All uncertainty intervals in this paper represent 95\% Confidence Intervals.},
and a concept-specific uncertainty of \(\sigma_C =\) 0.29 {[}0.1,
0.49{]}. Adults had a label-specific uncertainty of \(\sigma_S =\) 0.16
{[}0.05, 0.28{]}, and a concept-specific uncertainty of \(\sigma_C =\)
0.14 {[}0.05, 0.23{]}. As predicted, the uncertainty parameters were
larger for children than they were for adults.

\section{General Discussion}\label{general-discussion}

This paper studies the hypothesis that some seemingly stage-like
patterns in cognitive development can be characterized in a continuous
fashion. We used as a case study the seminal work of Stager \& Werker
(1997) showing a discrepancy between children's speech perception
abilities and their word learning skills. While most previous work
debated the source of this discrepancy, here we were more interested in
its mechanism of development.

Building on some previous findings and discussions (e.g., Swingley,
2007; Yoshida et al., 2009), we proposed a model which assumes that
subtle perceptual features are correctly encoded during word meaning
learning, but that the precision of this encoding improves across
development. We tested the model's predictions against data collected
from preschool children and adults and we showed that development can be
characterized as a continuous refinement in qualitatively similar
representations across the life span.

The model made a new prediction which we tested experimentally for the
first time: learning similar words is not only modulated by the
similarity of their phonological forms, but also by the visual
similarity of their semantic referents. More generally, since visual
similarity is an early organizing feature in the semantic domain (e.g.,
Wojcik \& Saffran, 2013), our finding suggests that children may
prioritize the acquisition of words that are quite distant in the
semantic space. This suggestion is supported by recent findings based on
the investigation of early vocabulary growth (Engelthaler \& Hills,
2017; Sizemore, Karuza, Giusti, \& Bassett, 2018). That said, further
work is needed to explore the effect on word learning of other semantic
dimensions that could be encoded by children (e.g.,
conceptual/functional features).

One limitation of this work is that the model was fit to data from
children at a relatively older age (4-5 years old) than what is
typically studied in the literature (around 14 month-old). While an
important conclusion of this paper is that the same process may apply to
different developmental stages, it would be interesting to
experimentally verify the generality of this conclusion in the context
of early language acquisition. A challenge to this next step would be to
construct a computational framework which would compare performance
across heterogeneous testing methods (e.g., looking time
vs.~forced-choice).

To sum, this paper proposes a model that accounts for the development of
an important aspect of word learning. The findings show that data can be
explained based on a continuous process operating over similar
representations across development. We used a case from word learning as
an example, but the same idea might apply to other aspects of cognitive
development which are tyipcally thought of as stage-like (e.g.,
acquisition of a theory of mind). A crucial task would be to investigate
the extent to which such discontinuities emerge due to genuine
qualitative changes, and the extent to which they reflect the
granularity of the researchers' own measurement tools.

\section{References}\label{references}

\setlength{\parindent}{-0.1in} \setlength{\leftskip}{0.125in} \noindent

\hypertarget{refs}{}
\hypertarget{ref-barr2013}{}
Barr, D., Levy, R., Scheepers, C., \& Tily, H. (2013). Random effects
structure for confirmatory hypothesis testing: Keep it maximal.
\emph{Journal of Memory and Language}, \emph{68}(3).

\hypertarget{ref-carey2009}{}
Carey, S. (2009). \emph{The origin of concepts}. Oxford University
Press.

\hypertarget{ref-dutoit1996}{}
Dutoit, T., Pagel, V., Pierret, N., Bataille, F., \& Van der Vrecken, O.
(1996). The mbrola project: Towards a set of high quality speech
synthesizers free of use for non commercial purposes. In
\emph{Proceedings of ICSLP} (Vol. 3). IEEE.

\hypertarget{ref-engelthaler2017}{}
Engelthaler, T., \& Hills, T. T. (2017). Feature biases in early word
learning: Network distinctiveness predicts age of acquisition.
\emph{Cognitive Science}, \emph{41}.

\hypertarget{ref-fennell2010}{}
Fennell, C., \& Waxman, S. (2010). What paradox? Referential cues allow
for infant use of phonetic detail in word learning. \emph{Child
Development}, \emph{81}.

\hypertarget{ref-hofer2017}{}
Hofer, M., \& Levy, R. (2017). Modeling Sources of Uncertainty in Spoken
Word Learning. In \emph{Proceedings of the 39th Annual Meeting of the
Cognitive Science Society}.

\hypertarget{ref-lewis2013}{}
Lewis, M., \& Frank, M. (2013). An integrated model of concept learning
and word-concept mapping. In \emph{Proceedings of the annual meeting of
the cognitive science society} (Vol. 35).

\hypertarget{ref-pajak2016}{}
Pajak, B., Creel, S., \& Levy, R. (2016). Difficulty in learning
similar-sounding words: A developmental stage or a general property of
learning? \emph{Journal of Experimental Psychology: Learning, Memory,
and Cognition}, \emph{42}(9).

\hypertarget{ref-piaget1954}{}
Piaget, J. (1954). \emph{The construction of reality in the child}. New
York, NY, US: Basic Books.

\hypertarget{ref-rost2009}{}
Rost, G., \& McMurray, B. (2009). Speaker variability augments
phonological processing in early word learning. \emph{Developmental
Science}, \emph{12}.

\hypertarget{ref-sizemore2018}{}
Sizemore, A. E., Karuza, E. A., Giusti, C., \& Bassett, D. S. (2018).
Knowledge gaps in the early growth of semantic feature networks.
\emph{Nature Human Behaviour}, \emph{2}(9).

\hypertarget{ref-stager1997}{}
Stager, C., \& Werker, J. (1997). Infants listen for more phonetic
detail in speech perception than in word-learning tasks. \emph{Nature},
\emph{388}(6640).

\hypertarget{ref-swingley2007}{}
Swingley, D. (2007). Lexical exposure and word-form encoding in
1.5-year-olds. \emph{Developmental Psychology}, \emph{43}(2).

\hypertarget{ref-thiessen2007}{}
Thiessen, E. (2007). The effect of distributional information on
children's use of phonemic contrasts. \emph{Journal of Memory and
Language}, \emph{56}.

\hypertarget{ref-werker2002}{}
Werker, J., Fennell, C., Corcoran, K., \& Stager, C. (2002). Infants'
ability to learn phonetically similar words: Effects of age and
vocabulary size. \emph{Infancy}, \emph{3}.

\hypertarget{ref-white2013}{}
White, K., Yee, E., Blumstein, S., \& Morgan, J. (2013). Adults show
less sensitivity to phonetic detail in unfamiliar words, too.
\emph{Journal of Memory and Language}, \emph{68}(4).

\hypertarget{ref-wojcik2013}{}
Wojcik, E., \& Saffran, J. (2013). The ontogeny of lexical networks:
Toddlers encode the relationships among referents when learning novel
words. \emph{Psychological Science}, \emph{24}(10).

\hypertarget{ref-yoshida2009}{}
Yoshida, K., Fennell, C., Swingley, D., \& Werker, J. (2009).
14-month-olds learn similar-sounding words. \emph{Developmental
Science}, \emph{12}.

\end{document}