### 2019-05-14 Analysis/process.Rmd
cogsci camera ready

---
title: "WordRec Experiment"
date: "January 13, 2017"
output:
  html_document:
    number_sections: yes
    toc: yes
---

Libraries.

```{r}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(langcog)
library(ggthemes)
library(lme4)
library(nlstools)
library(MuMIn)
library(purrr)
theme_set(theme_few())
```

Sample size:
The following are the analysese done with adult pilot data.
We plan to collect data from 100 participants on Amazaon MTurk.

Adult data.

```{r}

#The following is a test with adult pilot data
#d_adult <- read_delim("adults_final.txt", delim = " ")  %>%
d_adult <- read_delim("adults_100.txt", delim = " ")  %>%
  filter(type == "Test", #Here I will  filter out particiapants whose first language is arabic or hindi
         other != "yes") %>%
  mutate(iscorrect=ifelse(answer==correct, 1, 0))

d_adult_score <- d_adult %>%
  group_by(ID, sound_dist, concept_dist) %>%
  dplyr::summarise(mean = mean(iscorrect))

d_adult_excl <- d_adult_score %>%
  filter(sound_dist == 'catch') %>%
  ungroup() %>%
  select(ID, mean) %>%
  dplyr::rename(score = mean)
  #select(ID, score)

d_adult_good <- d_adult %>%
  left_join(d_adult_excl) %>%
  filter(score == 1) %>%
  filter(sound_dist != 'catch')

feather::write_feather(d_adult_good, "../saved_data/data_adult.feather")
```

Children data

```{r}
d_child <- read_delim("kids_all.txt", delim = " ") %>%
  filter(type == "Test") %>%
  filter(age != 6) %>%
  filter(code != 'test',
         !is.na(code)) %>%
  filter(grepl('soph', code))%>%
  mutate(iscorrect=ifelse(answer==correct, 1, 0))

d_child_score <- d_child %>%
  group_by(ID, sound_dist, concept_dist) %>%
  dplyr::summarise(mean = mean(iscorrect))

#Catch
d_child_excl <- d_child_score %>%
  filter(sound_dist == 'catch') %>%
  ungroup() %>%
  select(ID, mean) %>%
  dplyr::rename(score = mean)

d_child_good <- d_child %>%
  left_join(d_child_excl) %>%
  filter(score > 0.5) %>%
  filter(sound_dist != 'catch')

feather::write_feather(d_child_good, "../saved_data/data_child.feather")

```

Similarity data

```{r}

d_sim <- read_delim("sim_results.txt", delim = " ")  %>%
  filter(condition != "obj_prac" & condition != "sound_prac") %>%
  filter(problem == 'No') %>%
  mutate(scaled = (answer-1)/6)

d_sim_summary <- d_sim %>%
  group_by(condition, distance) %>%
  summarise(mean = mean(scaled),
                   sd = sd(scaled),
                   n = n()) %>%
  mutate(se = sd / sqrt(n),
         lower = mean - qt(1 - (0.05 / 2), n - 1) * se,
         upper = mean + qt(1 - (0.05 / 2), n - 1) * se) %>%
  select(-sd, -n, -se)

d_sim_process <- d_sim_summary %>%
  bind_rows(data.frame(condition="obj_real", distance="1", mean= 1, lower=1,upper=1)) %>%
  mutate(stimuli = ifelse(condition =='obj_real','objects','sounds')) %>%
  filter(distance != "iden",
         distance != "0.6") %>%
  mutate(labels = ifelse(distance=='0.1','similar', 
                         ifelse(distance=='0.3','intermediate', 
                                ifelse(distance=='sim','similar',
                                       ifelse(distance=='med','intermediate', 'different'))))) %>%
  select(-condition, -distance)

d_sim_process$labels <- factor(d_sim_process$labels, levels = c("similar","intermediate","different"))

feather::write_feather(d_sim_process, "../saved_data/d_sim_process.feather")

#d_sim_transform <-  d_sim_summary %>%
#  mutate(normalized =ifelse(stimuli=='sounds', mean/0.926, mean))  #Here I only normalize sounds, objects are already normalized

ggplot(d_sim_process, 
      aes(x = labels, y = mean, group = factor(stimuli), col=factor(stimuli))) +
  #geom_point(size=2)+
  geom_line()+
  geom_pointrange(aes(ymin = lower, ymax = upper), 
                 position = position_dodge(width = .1),
                 size=0.2) + 
  scale_colour_discrete(name="Stimuli")+
  theme_few()+
theme(legend.title = element_text(size=11),
      legend.text=element_text(size=11),
      axis.text = element_text(size = 11),
      aspect.ratio = 0.7
      ) +
  ylim(c(0,1.1)) +
  xlab("Stimuli group")+ ylab("Perceptual distance (experimental)")

```

Summary children
```{r}
sum_child <- d_child_good %>%
  group_by(sound_dist, concept_dist) %>%
  dplyr::summarise(mean = mean(iscorrect),
                   sd = sd(iscorrect),
                   n = n()) %>%
  mutate(se = sd / sqrt(n),
         lower = mean - qt(1 - (0.05 / 2), n - 1) * se,
         upper = mean + qt(1 - (0.05 / 2), n - 1) * se) %>%
  select(-sd, -n, -se)

```

Plot children
```{r}

sum_child$concept_dist <- factor(sum_child$concept_dist, levels = c("similar", "different"))
sum_child$sound_dist <- factor(sum_child$sound_dist, levels = c("similar", "different"))

plot_child <- ggplot(sum_child, 
      aes(x = sound_dist, y = mean, group = factor(concept_dist), col = factor(concept_dist))) +
  #geom_point()+
  geom_pointrange(aes(ymin = lower, ymax = upper), 
                 position = position_dodge(width = .1)) + 
  geom_line() + 
  geom_hline(yintercept = 0.50, linetype="dashed") +
  xlab("Auditory similary") +ylab("Accuracy")+
  scale_colour_discrete(name="Visual similarity")+
  theme_few()+
theme(legend.title = element_text(size=11),
      legend.text=element_text(size=11),
      axis.text = element_text(size = 11)
      )# + facet_grid(.~block)

plot_child

```
Summary adults

```{r}

sum_adult <- d_adult_good %>%
  group_by(sound_dist, concept_dist) %>%
  dplyr::summarise(mean = mean(iscorrect),
                   sd = sd(iscorrect),
                   n = n()) %>%
  mutate(se = sd / sqrt(n),
         lower.ci = mean - qt(1 - (0.05 / 2), n - 1) * se,
         upper.ci = mean + qt(1 - (0.05 / 2), n - 1) * se) %>%
  select(-sd, -n, -se)

```

```{r}

sum_adult$concept_dist <- factor(sum_adult$concept_dist, levels = c("similar", "different"))
sum_adult$sound_dist <- factor(sum_adult$sound_dist, levels = c("similar", "different"))

plot_adult <- ggplot(sum_adult, 
       aes(x = sound_dist, y = mean, group = factor(concept_dist), col = factor(concept_dist))) +

  geom_pointrange(aes(ymin = lower.ci, ymax = upper.ci), 
                  position = position_dodge(width = .1)) +

  geom_hline(yintercept = .50, linetype="dashed") + 
  geom_line() +
  xlab("Auditory similary") +ylab("Accuracy")+
  #scale_colour_manual(values = c("red", "blue"), name="Vis Dist")+
  scale_colour_discrete(name="Visual similarity")+

  #scale_y_continuous(limits = c(0.45, 0.9))+
  theme_few()+
theme(legend.title = element_text(size=11),
      legend.text=element_text(size=11),
      axis.text = element_text(size = 11)
      ) #+ facet_grid(. ~ block)

#theme(plot.title = element_text(size = 12, face = "bold"),
#    legend.title=element_text(size=10), 
#    legend.text=element_text(size=9))
plot_adult

```

Regression Model
```{r}

#Code and center levels
data_ch <- d_child_good %>%
  mutate(s_dist=ifelse(sound_dist=="similar", -1,1),
         c_dist=ifelse(concept_dist=="similar", -1, 1))

data_ad <- d_adult_good %>%
  mutate(s_dist=ifelse(sound_dist=="similar", -1,1),
         c_dist=ifelse(concept_dist=="similar", -1, 1))

#Add a centerd variable for age group and combine both data 
data_both_ch <- data_ch %>%
  select(ID, c_dist, s_dist, item_object, item_sound, iscorrect) %>%
  mutate(age = -1)

data_both_ad <- data_ad %>%
  select(ID, c_dist, s_dist, concept, sound,iscorrect) %>%
  rename(item_object = concept,
         item_sound = sound) %>%
  mutate(age = 1)

data_both <- data_both_ch %>%
  bind_rows(data_both_ad)

#Maximal random structure
model_ch <- glmer(iscorrect ~ s_dist * c_dist + 
                 (s_dist * c_dist  | ID) + 
                 (c_dist | item_object) + 
                 (s_dist | item_sound), 
               data = data_ch, family = binomial(),
               glmerControl(optimizer="bobyqa", optCtrl = list(maxfun = 100000)))

model_ad <- glmer(iscorrect ~ s_dist * c_dist + 
                 (s_dist * c_dist  | ID) + 
                 (c_dist | sound) + 
                 (s_dist | concept), 
               data = data_ad, family = binomial(),
               glmerControl(optimizer="bobyqa", optCtrl = list(maxfun = 100000)))

model_both <- glmer(iscorrect ~ s_dist * c_dist * age+ 
                 (s_dist * c_dist  | ID) + 
                 (c_dist | item_object) + 
                 (s_dist | item_sound) +
                 (s_dist * c_dist | age), 
               data = data_both, family = binomial(),
               glmerControl(optimizer="bobyqa", optCtrl = list(maxfun = 100000)))

stargazer(model_ch, model_ad, 
          column.labels = c("Children","Adults"),
          model.numbers=FALSE,
          intercept.bottom = FALSE,
          covariate.labels=c("(Intercept)","Sound ","Object","Sound x Object"),
          no.space=TRUE,
          single.row=TRUE,
          star.cutoffs =c(0.05, 0.01, 0.001),
          digits=2,
          omit.stat = "all",
          table.layout ="-c-!t-n",
          header=FALSE,
          #notes=c("[0.*]","[0.**]"),
          title="Estimates of predictor coefficients (and their standard errors) by age group in the regression model")

```

Use values of perceptual distance from the similarity judgement task

```{r}
#Process the data to prepare for the fit

#Extract distance values
dist_sim_c <- round(d_sim_process$mean[which(d_sim_process$condition=='obj_real' & d_sim_process$labels=='similar')],2)
dist_inter_c <- round(d_sim_process$mean[which(d_sim_process$condition=='obj_real' & d_sim_process$labels=='intermediate')],2)
dist_diff_c <- round(d_sim_process$mean[which(d_sim_process$condition=='obj_real' & d_sim_process$labels=='different')],2)

dist_sim_s <- d_sim_process$mean[which(d_sim_process$condition=='sound_real' & d_sim_process$labels=='similar')]
dist_inter_s <- d_sim_process$mean[which(d_sim_process$condition=='sound_real' & d_sim_process$labels=='intermediate')]
dist_diff_s <- d_sim_process$mean[which(d_sim_process$condition=='sound_real' & d_sim_process$labels=='different')]

#Here normalize the sound with respect to the most distant pair

dist_sim_s <- round(dist_sim_s/dist_diff_s,2)
dist_inter_s <- round(dist_inter_s/dist_diff_s,2)
dist_diff_s <- round(dist_diff_s/dist_diff_s,2)

#For adults: use the similar vs. intermediate pairs
data_fit_adult <- sum_adult %>%
  rename(lower  = lower.ci,
         upper = upper.ci) %>%
  mutate(group = 'adults',
         s_dist = ifelse(sound_dist =='different',  dist_inter_s, dist_sim_s),
         c_dist = ifelse(concept_dist =='different', dist_inter_c, dist_sim_c))

#For children: use the intermediate vs. different pairs 
data_fit_child <- sum_child %>% 
   mutate(group = 'children',
         s_dist = ifelse(sound_dist =='different',  dist_diff_s, dist_inter_s),
         c_dist = ifelse(concept_dist =='different', dist_diff_c, dist_inter_c))

data_fit <- data_fit_adult %>%
  bind_rows(data_fit_child)
```

Now I need to predict the mean accuracy as a function of sound and meaning similarity 
```{r}

#This the analytical solution of the posterior + luce rule
combine_cues <- function(s, o, Vr) {
  s1 = 0
  o1 = 0
  no_acc <- exp(-(s1-s)^2/(2*(Vr)^2)) + exp(-(o1-o)^2/(2*(Vr)^2))
  acc <- 1 + exp(-((s1-s)^2+(o1-o)^2)/(2*(Vr)^2)) 
  acc_norm <- acc/(acc+no_acc)

  return(acc_norm)
}

no_combine_cues <- function(s, o, Vr) {
  s1 = 0
  o1 = 0

   no_acc <- exp(-(s1-s)^2/(2*(Vr)^2)) 
  acc <- 1 
  acc_norm <- acc/(acc+no_acc)

  return(acc_norm)
}

combine_cues_2 <- function(s, o, Vr_s, Vr_o) {
  s1 = 0
  o1 = 0
  no_acc <- exp(-(s1-s)^2/(2*(Vr_s)^2)) + exp(-(o1-o)^2/(2*(Vr_o)^2))
  acc <- 1 + exp(-((s1-s)^2/(2*(Vr_s)^2)+(o1-o)^2/(2*(Vr_o)^2))) 
  acc_norm <- acc/(acc+no_acc)

  return(acc_norm)
}

fit_adults <- nls(mean ~ combine_cues(s_dist, c_dist, Vr), data=subset(data_fit, group=='adults'), start=list(Vr=0.1), lower=c(0), algorithm="port")

fit_adults_no <- nls(mean ~ no_combine_cues(s_dist, c_dist, Vr), data=subset(data_fit, group=='adults'), start=list(Vr=0.5), lower=c(0), algorithm="port")

fit_adults_2 <- nls(mean ~ combine_cues_2(s_dist, c_dist, Vr_s, Vr_o), data=subset(data_fit, group=='adults'), start=list(Vr_s=0.1, Vr_o=0.1), lower=c(0,0), algorithm="port")

Vr_ad <- coef(fit_adults)
Vr_ad_no <- coef(fit_adults_no)
Vr_s_ad <- coef(fit_adults_2)[1]
Vr_o_ad <- coef(fit_adults_2)[2]

fit_children <- nls(mean ~ combine_cues(s_dist, c_dist, Vr), data=subset(data_fit, group=='children'), start=list(Vr=0.5), lower=c(0), algorithm="port")

fit_children_no <- nls(mean ~ no_combine_cues(s_dist, c_dist, Vr), data=subset(data_fit, group=='children'), start=list(Vr=0.5), lower=c(0), algorithm="port")

fit_children_2 <- nls(mean ~ combine_cues_2(s_dist, c_dist, Vr_s, Vr_o), data=subset(data_fit, group=='children'), start=list(Vr_s=0.5, Vr_o=0.3), lower=c(0,0), algorithm="port")

Vr_ch <- coef(fit_children)
Vr_ch_no <- coef(fit_children_no)
Vr_s_ch <- coef(fit_children_2)[1]
Vr_o_ch <- coef(fit_children_2)[2]

model1_val <- as.data.frame(confint2(fit_children)) %>%
  mutate(val= Vr_ch, 
         model=1, 
         group = "children") %>%
  bind_rows(as.data.frame(confint2(fit_adults)) %>%
              mutate(val= Vr_ad, 
                     model=1, 
                     group = "adults"))

model_no_val <- as.data.frame(confint2(fit_children_no)) %>%
  mutate(val= Vr_ch_no, 
         model=1, 
         group = "children") %>%
  bind_rows(as.data.frame(confint2(fit_adults)) %>%
              mutate(val= Vr_ad_no, 
                     model=1, 
                     group = "adults"))

model2_val <- tibble::rownames_to_column(as.data.frame(confint2(fit_children_2)), "param") %>%
  mutate(val= ifelse(param=='Vr_s',Vr_s_ch, Vr_o_ch),
         model=2, 
         group = "children") %>%
  bind_rows(tibble::rownames_to_column(as.data.frame(confint2(fit_adults_2)), "param") %>%
              mutate(val= ifelse(param=='Vr_s',Vr_s_ad, Vr_o_ad),
                     model=2, 
                     group = "adults"))

#Model comparison

#model_comparison <- tibble::rownames_to_column(AIC(fit_adults, fit_adults_2), 'model') %>%
#  bind_rows(tibble::rownames_to_column(AIC(fit_children, fit_children_2), 'model'))

feather::write_feather(model1_val, "../saved_data/values_mod1.feather")  
feather::write_feather(model_no_val, "../saved_data/values_mod_no.feather")

feather::write_feather(model2_val, "../saved_data/values_mod2.feather")

#feather::write_feather(model_comparison, "../saved_data/model_comparison.feather")

model2_val <- c(confint(fit_children_2), Vr_s_ch, Vr_o_ch, model="2", group = "children") %>%
  bind_rows(c(confint(fit_adults_2), Vr_s_ad, Vr_o_ad, model="2", group = "adults"))

model_data <- data_fit %>%
  mutate(prediction = ifelse(group=='adults', 
                             combine_cues(s_dist, c_dist, Vr_ad),
                             combine_cues(s_dist, c_dist, Vr_ch)
                             )) %>%
  mutate(prediction_no = ifelse(group=='adults', 
                             no_combine_cues(s_dist, c_dist, Vr_ad_no),
                             no_combine_cues(s_dist, c_dist, Vr_ch_no)
                             )) %>%
  mutate(prediction_2 = ifelse(group=='adults', 
                             combine_cues_2(s_dist, c_dist, Vr_s_ad, Vr_o_ad),
                             combine_cues_2(s_dist, c_dist, Vr_s_ch, Vr_o_ch)
                             )) %>%
  rename(human = mean)

feather::write_feather(model_data, "../saved_data/data_correlations.feather")
R2_model1 <-  round(cor(model_data$human, model_data$prediction)^2, 2)
R2_model_no <-  round(cor(model_data$human, model_data$prediction_no)^2, 2)
R2_model2 <-  round(cor(model_data$human, model_data$prediction_2)^2, 2)

model_data_plot <-  model_data %>%
  gather(key, value, prediction, prediction_no, prediction_2) #%>%
  #mutate(ci.upper = ifelse(key=='human', upper, value),
  #       ci.lower = ifelse(key=='human', lower, value))

feather::write_feather(model_data_plot, "../saved_data/data_human_model.feather")

```

```{r}

model_data_plot$group <- factor(model_data_plot$group, levels = c("children", "adults"))

ggplot(subset(model_data_plot, key == "human"), 
      aes(x = sound_dist, y = value, 
          group = factor(concept_dist), 
          col = factor(concept_dist)
          )) +
  #geom_point()+
  geom_pointrange(aes(ymin = ci.lower, ymax = ci.upper), 
                 position = position_dodge(width = .1),
                 size=0.2) + 
  geom_line() + 
  geom_line(data=subset(model_data_plot, key == "prediction"), linetype=2, size=0.5, col='black')+
  geom_point(data=subset(model_data_plot, key == "prediction"), size=1, col='black')+
  geom_hline(yintercept = 0.50, linetype="dashed") +
  xlab("Auditory similary") +ylab("Accuracy")+
  scale_colour_discrete(name="Visual similarity")+
  theme_few()+
theme(legend.title = element_text(size=11),
      legend.text=element_text(size=11),
      axis.text = element_text(size = 11)
      ) + facet_grid(.~group) +
  theme(aspect.ratio = 0.7)

ggplot(subset(model_data_plot, key == "human"), 
      aes(x = sound_dist, y = value, 
          group = factor(concept_dist), 
          col = factor(concept_dist)
          )) +
  #geom_point()+
  geom_pointrange(aes(ymin = ci.lower, ymax = ci.upper), 
                 position = position_dodge(width = .1),
                 size=0.2) + 
  geom_line() + 
  geom_line(data=subset(model_data_plot, key == "prediction_2"), linetype=2, size=0.5, col='black')+
  geom_point(data=subset(model_data_plot, key == "prediction_2"), size=1, col='black')+
  geom_hline(yintercept = 0.50, linetype="dashed") +
  xlab("Auditory similary") +ylab("Accuracy")+
  scale_colour_discrete(name="Visual similarity")+
  theme_few()+
theme(legend.title = element_text(size=11),
      legend.text=element_text(size=11),
      axis.text = element_text(size = 11)
      ) + facet_grid(.~group) +
  theme(aspect.ratio = 0.7)

```

Simulations
```{r}

simulation1 <- function(s,o,Vr) {
  s1 = 0
  o1 = 0
  no_acc <- exp(-(s1-s)^2/(2*Vr)) + exp(-(o1-o)^2/(2*Vr))
  acc <- 1 + exp(-((s1-s)^2+(o1-o)^2)/(2*Vr)) 
  acc_norm <- acc/(acc+no_acc)

  return(acc_norm)
}

s <- seq(0,1,0.01)

simulation_data <- data.frame(x=s, y=simulation1(s, o=0.25, Vr=(0.15)^2), sigma = '0.15', object_distance = '0.25') %>%
  bind_rows(data.frame(x=s, y=simulation1(s, o=0.5, Vr=(0.15)^2), sigma = '0.15', object_distance = '0.5')) %>%
  bind_rows(data.frame(x=s, y=simulation1(s, o=1, Vr=(0.15)^2), sigma = '0.15', object_distance = '1')) %>%
  bind_rows(data.frame(x=s, y=simulation1(s, o=0.25, Vr=(0.3)^2), sigma = '0.3', object_distance = '0.25')) %>%
  bind_rows(data.frame(x=s, y=simulation1(s, o=0.5, Vr=(0.3)^2), sigma = '0.3', object_distance = '0.5')) %>%
  bind_rows(data.frame(x=s, y=simulation1(s, o=1, Vr=(0.3)^2), sigma = '0.3', object_distance = '1')) %>%
  bind_rows(data.frame(x=s, y=simulation1(s, o=0.25, Vr=(0.6)^2), sigma = '0.6', object_distance = '0.25')) %>%
  bind_rows(data.frame(x=s, y=simulation1(s, o=0.5, Vr=(0.6)^2), sigma = '0.6', object_distance = '0.5')) %>%
  bind_rows(data.frame(x=s, y=simulation1(s, o=1, Vr=(0.6)^2), sigma = '0.6', object_distance = '1'))

feather::write_feather(simulation_data, "../saved_data/simulations.feather")

ggplot(simulation_data, aes(x=x, y=y, col=sigma)) + 
  geom_line() + 
  theme_few()+
  theme(legend.title = element_text(size=11),
      legend.text=element_text(size=11),
      axis.text = element_text(size = 11),
      aspect.ratio = 0.7
      )+
  ylim(c(0.4,1))+
  geom_hline(yintercept = 0.50, linetype="dashed") +
  xlab("sound distance") + ylab("Prob. accurate") +
  facet_grid(.~object_distance)

```

plots 
```{r}

x <- 0

Gdist <- function(s0, Vr, x){

  return((1/sqrt(2*pi*Vr))*exp(-(s0-x)^2/(2*Vr)))
}

s = seq(-1.5,4,0.01)
y = Gdist(0,1,x)

illustration <- data.frame(x=s, y=Gdist(s0=0, Vr=1, s), rep = "1", case = "low accuracy", type='Disambiguation') %>%
  bind_rows(data.frame(x=s, y=Gdist(s0=2, Vr=1, s), rep = "2", case = "low accuracy", type='Disambiguation')) %>%
  bind_rows(data.frame(x=s, y=Gdist(s0=1, Vr=0.5, s), rep = "1", case = "high accuracy", type='Disambiguation')) %>%
  bind_rows(data.frame(x=s, y=Gdist(s0=2, Vr=0.5, s), rep = "2", case = "high accuracy", type='Disambiguation')) %>%
  bind_rows(data.frame(x=s, y=Gdist(s0=1, Vr=0.5, s), rep = "1", case = "low accuracy", type='Development')) %>%
  bind_rows(data.frame(x=s, y=Gdist(s0=1, Vr=0.5, s), rep = "2", case = "low accuracy", type='Development')) %>%
  bind_rows(data.frame(x=s, y=Gdist(s0=0, Vr=0.1, s), rep = "1", case = "high accuracy", type='Development')) %>%
  bind_rows(data.frame(x=s, y=Gdist(s0=2, Vr=0.1, s), rep = "2", case = "high accuracy", type='Development'))

illustration$case <- factor(illustration$case, levels = c("low accuracy", "high accuracy"))
illustration$type <- factor(illustration$type, levels = c("Disambiguation", "Development"))

probs <- subset(subset(illustration, x==0)) %>%
  group_by(type, case) %>%
  mutate(prob = round(y/sum(y),2)) %>%
  filter(rep=='1')

ggplot(illustration, aes(x=x, y=y, col=rep)) + 
  geom_line() + 
  theme_few()+
  theme(legend.title = element_text(size=11),
      legend.text=element_text(size=11),
      axis.text = element_text(size = 11),
      strip.text = element_text(size = 11),
      aspect.ratio = 0.5
      )+
  #geom_col(aes(x = x, y = y, fill = rep),
  #             position=position_dodge(), 
  #             width=0.2,
  #             data = subset(illustration, x==0))+
  #geom_text(aes(label=paste("Acc=", prob, sep="")),
  #          x=2, y=1, size=4, 
  #          fontface = "bold",
  #          data=probs)+
  xlab("Perceptual space") + ylab("Prob") +
  facet_grid(type~case) +
  theme(legend.position="none")
  #xlim(c(-1,3))

```

Regression

```{r}
data_ch <- d_child_good %>%
  mutate(s_dist=ifelse(sound_dist=="similar", -1,1),
         c_dist=ifelse(concept_dist=="similar", -1, 1)) %>%
  rename(accuracy = iscorrect)

data_ad <- d_adult_good %>%
  mutate(s_dist=ifelse(sound_dist=="similar", -1,1),
         c_dist=ifelse(concept_dist=="similar", -1, 1)) %>%
  rename(accuracy = iscorrect)

#Combine both  
data_both_ch <- data_ch %>%
  select(ID, c_dist, s_dist, item_object, item_sound, accuracy) %>%
  mutate(age = -1)

data_both_ad <- data_ad %>%
  select(ID, c_dist, s_dist, concept, sound, accuracy) %>%
  rename(item_object = concept,
         item_sound = sound) %>%
  mutate(age = 1) %>%
  mutate(ID2 = sub("+$", "_ad", ID)) %>%
  select(-ID) %>%
  rename(ID=ID2)

data_both <- data_both_ch %>%
  bind_rows(data_both_ad)

model_both <- glmer(accuracy ~ s_dist * c_dist * age+ 
                 (s_dist * c_dist  | ID) + 
                 (c_dist | item_object) + 
                 (s_dist | item_sound) +
                 (s_dist * c_dist | age), 
               data = data_both, family = binomial(),
               glmerControl(optimizer="bobyqa", optCtrl = list(maxfun = 100000)))

model_coef <- tibble::rownames_to_column(as.data.frame(coef(summary(model_both))), "effect")

feather::write_feather(model_coef, "../saved_data/model_regression.feather")

```