### 2019-07-22 journal/ms.tex

\documentclass[english,,man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Continuous developmental change can explain discontinuities in word learning},
            pdfauthor={Abdellah Fourtassi, Sophie Regan, \& Michael C. Frank},
            pdfkeywords={word learning, cognitive development, computational modeling},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

  \title{Continuous developmental change can explain discontinuities in word
learning}
    \author{Abdellah Fourtassi\textsuperscript{1}, Sophie Regan\textsuperscript{1},
\& Michael C. Frank\textsuperscript{1}}
    \date{}

\shorttitle{Continuous Development of Word Learning}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Department of Psychology, Stanford University}
\keywords{word learning, cognitive development, computational modeling}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

\usepackage{longtable}
\usepackage{lscape}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage[flushleft]{threeparttable}
\usepackage{threeparttablex}

\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

\usepackage{lineno}

\linenumbers
\usepackage{tipa}
\usepackage[sortcites=false,sorting=none]{biblatex}

\authornote{ Abdellah Fourtassi

Department of Psychology

Stanford University

50 Serra Mall

Jordan Hall, Building 420

Stanford, CA 94301

Correspondence concerning this article should be addressed to Abdellah
Fourtassi, Postal address. E-mail:
\href{mailto:afourtas@stanford.edu}{\nolinkurl{afourtas@stanford.edu}}}

\abstract{
``Cognitive development is often characterized in term of
discontinuities, but these discontinuities can sometimes be apparent
rather than actual and can arise from continuous developmental change.
To explore this idea, we use as a case study the finding by Stager and
Werker (1997) that children's early ability to distinguish similar
sounds does not automatically translate into word learning skills. Early
explanations proposed that children may not be able to encode subtle
phonetic contrasts when learning novel word meanings, thus suggesting a
discontinuous/stage-like pattern of development. However, later work has
revealed (e.g., through using simpler testing methods) that children do
encode such contrasts, thus favoring a continuous pattern of
development. Here we propose a probabilistic model describing how
development may proceed in a continuous fashion across the lifespan. The
model accounts for previously documented facts and provides new
predictions. We collected data from preschool children and adults, and
we showed that the model can explain various patterns of learning both
within the same age and across development. The findings suggest that
major aspects of cognitive development that are typically thought of as
discontinuities may emerge from simpler, continuous mechanisms.''

}

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{definition}
\newtheorem{example}{Example}[section]
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}[section]
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}
\maketitle

\section{Introduction}\label{introduction}

Cognitive development is sometimes characterized in terms of a
succession of discontinuous stages (Piaget, 1954). Although intuitively
appealing, stage theories can be challenging to integrate with theories
of learning, which typically posit that knowledge and skills improve
incrementally with experience. Indeed, one of the central challenges of
cognitive development has been to explain transitions between stages
which appear to be qualitatively different (Carey, 2009).

Nevertheless, at least in some cases, development may only appear to be
stage-like. This appearance can be due, for example, to the use of a
cognitively-demanding task which may mask learning, or to the use of
statistical thresholding (in particular, p-value \textless{} 0.05) which
can create a spurious dichotomy between success and failure in observing
a given behavior. In such cases, positing discontinuous stages is
unnecessary. Instead, a continuous model---involving similar
representations across the lifespan---may provide a simpler and more
transparent account of development.

We use a case study from word learning literature. Stager and Werker
(1997) first showed that children's early ability to distinguish similar
sounds does not automatically translate into word learning skills. The
authors measured word learning using an audio-visual habituation Switch
task. First, infants are familiarized with two word-object pairings
(e.g., Word A with Object A and Word B with Object B). Second, they are
tested using two types of trials. The control \enquote{same} trial
consists of a correct pairing (e.g., Word A with Object A) and the
\enquote{switch} trial consists of a wrong pairing (e.g., Word A with
Object B). If babies have correctly learned the association during the
familiarization, they are supposed to be surprised by the
\enquote{switch} trial and not by the \enquote{same} trial. The former
should thus result in a greater looking time compared to the latter
(Werker, Cohen, Lloyd, Casasola, \& Stager, 1998).

Though infants around 14-month old can distinguish perceptually similar
sound pairs such as \enquote{dih} and \enquote{bih}, they appear to fail
in mapping this pair to two different objects in the switch task. This
failure has initially been taken as evidence that 14-month olds do not
encode subtle sounds during meaning learning (Pater, Stager, \& Werker,
2004; Stager \& Werker, 1997). This interpretation suggested a
discontinuous/stage-like pattern of development whereby younger children
fail to encode the contrastive phonetic detail, whereas older children,
around 17 months (Werker, Fennell, Corcoran, \& Stager, 2002), typically
do.

The initial discontinuous interpretation has been challenged by
subsequent work. For instance, Yoshida, Fennell, Swingley, and Werker
(2009) investigated whether failure in the switch task reflects a lack
of sound encoding during \emph{familiarization}, or whether it is only
due to the demands of the \emph{testing} method which does not allow
learning below a certain threshold to be detected. They used the same
familiarization procedure as Stager and Werker (1997), but instead of
comparing the looking times in \enquote{same} and \enquote{switch}
trials, they tested infants using a two-alternative choice task
comparing fixations to target and distractor objects (Fernald, Perfors,
\& Marchman, 2006; Golinkoff, Hirsh-Pasek, Cauley, \& Gordon, 1987).
Using this simpler and finer-grained task, the researcher found evidence
for learning even in 14-month olds.

Another challenge to the discontinuous account of development came from
adult studies. If the mismatch between sound discrimination and word
learning is only a stage in early infancy, then this mismatch should
disappear by adulthood. Nonetheless, even adults show patterns of
learning that mirror those shown by 14-month-olds when the sound
contrasts more challenging (Pajak, Creel, \& Levy, 2016; White, Yee,
Blumstein, \& Morgan, 2013).

Some researchers (Pajak et al., 2016; e.g., Swingley, 2007; Yoshida et
al., 2009) proposed that the phonological form may not be encoded in a
binary fashion, i.e., it is not the case that children either succeed or
fail in encoding minimal contrast when learning the meanings. Rather,
they may be encoding the phonological form of words in a probabilistic
fashion. According to this view, development does not so much involve a
qualitative shift (i.e., a sudden emergence of an ability that did not
exist before) as much as it consists in the continuous refinement of
initially noisy representations.

In a probabilistic account, a word can be represented as a probability
distribution over sound instances organized in a similarity space. The
probability is highest at the most typical sound instance. It decreases
as the instance becomes less typical. The precision of the
representation can be characterized by how much it tolerates slightly
atypical pronunciations. This tolerance is captured formally by the
variance of the probability distribution: larger variance indicates
higher tolerance and lower precision, whereas smaller variance indicates
lower tolerance and higher precision.

This framework can explain several findings. In Stager and Werker's
original experiment, children are supposed to associate one label
\enquote{bih} with object 1 and a second label \enquote{dih} with object
2. Though infants could learn that the label \enquote{bih} is a better
match to object 1 than \enquote{dih}, they could still judge the sound
\enquote{dih} as a plausible instance of the label \enquote{bih}, thanks
to the relatively large variance of the early encoding, and this
confusion leads to \enquote{failure} in the recognition task (Figure
\ref{fig:illus}, top). Though learning is small and is easily disrupted
by the Switch task, it can still be detected when less demanding methods
are used (Yoshida et al., 2009).

The learning accuracy increases (and is detected even by demanding tests
such as the Switch) for more distinct word-forms (e.g., \enquote{lif}
vs. \enquote{neem}) where the perceptual distance is large relative to
the variance (Figure \ref{fig:illus}, left). Distinctiveness can be
enhanced even for minimally different sounds when other cues highlight
their difference (Dautriche, Swingley, \& Christophe, 2015; Rost \&
McMurray, 2009, 2010; Thiessen, 2007; Yeung \& Werker, 2009).

Finally, development can be understood as an increase in the precision
(i.e., a decrease in the variance) of the probabilistic representations
(Figure \ref{fig:illus}, right). Such an increase in precision renders
minimally different sounds less confusing. Importantly, a more precise
representation still has a non-zero variance --- Learning difficulties
can still be induced with challenging stimuli or in cognitively
demanding situations as was demonstrate in adults studies (Pajak et al.,
2016; White et al., 2013).

\subsection{This study}\label{this-study}

The probabilistic account has been put forward to explain patterns of
learning and development at the qualitative level. However, it is
crucial to have a precise computational instantiation of this account
which would allow us to \emph{quantify} its explanatory power. We could
find one previous work that attempted to provide such a computational
instantiation (Hofer and Levy, 2017). However, this previous work was
designed with the goal of reproducing the results of a specific study
(Pajak et al., 2016) which focused on explaining the mismatch between
speech perception and word learning in adults rather than on exploring
the mechanism of development.

The present work proposes a model of word-pair learning based on the
probabilistic account. We tested the ability of this model to both
\emph{explain} various findings in previous experiments in both children
and adults (e.g., the fact that similar words are harder to learn than
different words) and to \emph{predict} new learning patterns that have
not been tested before (i.e., the effect of the referents' similarity on
word learning). Crucially, we explored the extent to which development
can be understood as a continuous refinement in similar representations
across the lifespan.

The paper is organized as follows. First, we introduce the model and we
explain how it allows us to characterize behavior in a word learning
task which resembles the one used in Stager and Werker (1997) and
Yoshida et al. (2009). Then we explore the predictions of the model
through simulating its behavior across different parameter settings.
Next we quantify the extent to which the model's predictions account for
human data we collected from both preschool children and adults.
Finally, we discuss the results in the lights of existing accounts of
word development.

\begin{figure}

{\centering \includegraphics[width=300px]{figs/illustration}

}

\caption{An illustration of the probabilistic/continuous account using simulated data. A word is represented with a distribution over the perceptual space (indicated in red or blue). When the uncertainty of the representation is large relative to the distance between the stimuli (top panel), an instance of the red category (indicated with a star) could also be a plausible instance of the green category, hence the low recognition accuracy score. The accuracy increases when the stimuli are less similar (left panel), or when the representation are more precise (right panel).}\label{fig:illus}
\end{figure}

\section{Model}\label{model}

\subsection{Probabilistic structure}\label{probabilistic-structure}

Our model consists of a set of variables describing the general process
of spoken word recognition in a referential situation. These variables
are related in a way that reflects the simple generative scenario
represented graphically in Figure \ref{fig:model}. When a speaker utters
a sound in the presence of an object, the observer assumes that the
object \(o\) activated the concept \(C\) in the speaker's mind. The
concept prompted the corresponding label \(L\). Finally, the label was
physically instantiated by the sound \(s\).

\begin{figure}

{\centering \includegraphics[width=300px]{figs/model}

}

\caption{Graphical representation of our model. Circles indicate random variables (shading indicates observed variables). The squares indicate fixed model parameters.}\label{fig:model}
\end{figure}

A similar probabilistic structure was used by Lewis and Frank (2013) to
model concept learning, and by Hofer and Levy (2017) to model spoken
word learning. However, the first study assumed that the sounds are
heard unambiguously, and the second assumed the concepts are observed
unambiguously. In our model, we assume that both labels and concepts are
observed with a certain amount of perceptual noise, which we assume, for
simplicity, is captured by a normal distribution:

\[ p(o | C) \sim  \mathcal{N}(\mu_C, \sigma^2_C) \]

\[ p(s| L) \sim  \mathcal{N}(\mu_L, \sigma^2_L) \]

Finally, we assume there to be one-to-one mappings between concepts and
labels and that observers have successfully learned these mappings
during the exposure phase: \[
P(L_i|C_j) = 
\begin{cases}
  1 & \text{if  }  i=j \\  
  0  & \text{otherwise  }
\end{cases}
\]

\subsection{Inference}\label{inference}

The learner hears a sound \(s\) and has to decide which object \(o\)
provides an optimal match to this sound (see Figure \ref{fig:task}). To
this end, they must compute the probability \(P(o|s)\) for all possible
objects. This probability can be computed by summing over all possible
concepts and labels:
\[P(o|s)=\sum_{C,L} P(o, C, L| s) \propto \sum_{C,L} P(o, C, L, s) \]

The joint probability \(P(o, C, L, s)\) is obtained by factoring the
Bayesian network in Figure \ref{fig:model}:
\[P(o,C,L,s) = P(s|L)P(L|C)P(C|o)P(o) \]

which can be transformed using Bayes rule into:

\[P(o,C,L,s) = P(s|L)P(L|C)P(o|C)P(C) \]

Finally, assuming that the concepts' prior probability is uniformly
distributed\footnote{This is a reasonable assumption in our particular case given the similarity of the concepts used in each naming situation in our experiment.},
we obtain the following expression, where all conditional dependencies
are now well defined:

\[
P(o|s) = \frac{\sum_{C,L} P(s|L)P(o|C)P(L|C)}{\sum_{o} \sum_{C,L} P(s|L)P(o|C)P(L|C)}
\]

\subsection{Task and model
predictions}\label{task-and-model-predictions}

\begin{figure}[t]

{\centering \includegraphics[width=3.75in]{figs/task}

}

\caption{An overview of the task used in this study.}\label{fig:task}
\end{figure}

We use the model to predict word learning in a task similar to the one
introduced by Stager and Werker (1997). We used a modified version of
the task where the testing method consists in a two-alternative forced
choice (Yoshida et al., 2009). In this task, participants are first
exposed to the association between pairs of nonsense words (e.g.,
\enquote{lif}/\enquote{neem}) and pairs of objects. The word-object
associations are introduced sequentially. After this exposure phase,
participants perform a series of test trials. In each of these trials,
one of the two sounds is uttered (e.g., \enquote{lif}) and participants
choose the corresponding object from the two alternatives. An overview
of the task is shown in Figure \ref{fig:task}.

From the general expression (1), we derive three exact analytical
solutions instantiating different learning assumptions. The first
solution is derived by assuming that the labels are recovered from
sounds with a certain level of uncertainty \(\sigma_L\), but that
concepts are unambiguously recovered from the observed objects, i.e.,
\(\sigma_C \rightarrow 0\). This assumption has been made --- whether
implicitly or explicitly --- by most previous work in this line of
research. One important implication of this assumption is that only the
similarity of word sounds modulates success in word learning, not the
similarity of the referents (as long as these referents are
differentiated perceptually). This assumption yields the following
probability function:

\begin{equation}
P(o_T|s)= \frac{1}{1 + e^{-\frac{\Delta s^2}{2\sigma_L^2}}}
\end{equation}

The second solution is derived by making the more general assumption
that both the labels and the concepts are recovered with ambiguity from
the sounds and objects. We first introduce the simplifying assumption
that the label-related uncertainty \(\sigma_L\) and the concept-related
uncertainty \(\sigma_C\) are of a similar magnitude, i.e.,
\(\sigma_C \approx \sigma_L = \sigma\). This assumption makes the
prediction that the sound similarity and the object similarity impact
word learning accuracy in exactly the same way. Furthermore, it allows
us to study the behavior of the model with minimal free parameters.

\begin{equation}
P(o_T|s)= \frac{1 + e^{- \frac{\Delta s^2 + \Delta o^2}{2\sigma^2}}}{1 + e^{-\frac{\Delta s^2 + \Delta o^2}{2\sigma^2}}+ e^{-\frac{\Delta s^2}{2\sigma^2}} + e^{-\frac{\Delta o^2}{2\sigma^2}}}
\end{equation}

We finally derive the third (and most general) solution which allows
label- and concept-related uncertainties to vary independently.

\begin{equation}
P(o_T|s)= \frac{1 + e^{- (\frac{\Delta s^2}{2\sigma_L^2}+ \frac{\Delta o^2}{2\sigma_C^2})}}{1 + e^{-(\frac{\Delta s^2}{2\sigma_L^2}+ \frac{\Delta o^2}{2\sigma_C^2})}+ e^{-\frac{\Delta s^2}{2\sigma_L^2}} + e^{-\frac{\Delta o^2}{2\sigma_C^2}}}
\end{equation}

In order to understand the predictions of the models (especially the
more general ones, i.e., Model 2 and 3), Figure \ref{fig:simulation}
show some simulations of the accuracy \(P(o_T|s)\) as a function of the
distinctiveness parameters (\(\Delta s\) and \(\Delta o\)) and the
uncertainty parameters \(\sigma_L\) and \(\sigma_C\).

\begin{figure}[htbp]
\centering
\includegraphics{ms_files/figure-latex/simulation-1.pdf}
\caption{\label{fig:simulation}The predicted probability of accurate
responses in the testing phase as a function of stimuli distinctiveness
\(\Delta s\) and \(\Delta o\) and representation precision \(\sigma\)
(For clarity, we assume here that \(\sigma\)=\(\sigma_C\)=\(\sigma_L\)).
Dashed line represents chance.}
\end{figure}

The simulations explain two experimental results from previous studies
and make one new prediction:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\item
  For fixed values of \(\Delta o\) and \(\sigma\), the probability of
  accurate responses increases as a function of \(\Delta s\). This
  pattern accounts for the fact that similar sounds are generally more
  challenging to learn than different sounds for both children (Stager
  \& Werker, 1997) and adults (Pajak et al., 2016).
\item
  For fixed values of \(\Delta s\) and \(\Delta o\), accuracy increases
  when the representational uncertainty \(\sigma\) decreases. This fact
  may explain development, i.e., younger children have noisier
  representations (see Swingley, 2007; Yoshida et al., 2009), which
  leads to lower word recognition accuracy, especially for
  similar-sounding words.
\item
  For fixed values of \(\Delta s\) and \(\sigma\), accuracy increases
  with the visual distance between the semantic referents \(\Delta o\).
  This is a new prediction that our model makes. Previous work studied
  the effect of several bottom-up and top-down properties in
  disambiguating similar sounding words (e.g., Fennell \& Waxman, 2010;
  Rost \& McMurray, 2009; Thiessen, 2007), but to our knowledge, no
  previous study in the literature tested the effect of the visual
  distance between the semantic referents.
\end{enumerate}

\section{Experiment}\label{experiment}

In this experiment, we tested participants in the word learning task
introduced above (Figure \ref{fig:task}). More precisely, we explored
the predictions related to both distinctiveness and precision. Sound
similarity (\(\Delta s\)) and object similarity (\(\Delta o\)) were
varied simultaneously in a within-subject design. Two age groups
(preschool children and adults) were tested on the same task to explore
whether development can be characterized with the uncertainty
parameters, \(\sigma_C\) and \(\sigma_L\). The experiment, sample size,
exclusion criteria and the model's main predictions were
pre-registered.\footnote{https://osf.io/942gv/}

\subsection{Methods}\label{methods}

\subsubsection{Participants}\label{participants}

We report data from \(N=\) 63 children ages 4-5 years from the Bing
Nursery School on Stanford University's campus. An additional \(N=\) 39
children participated but were removed from analyses because they were
not above chance on the catch trials due to the challenging nature of
our procedure (see below). We also report data from \(N=\) 74 adult
participants tested on Amazon Mechanical Turk. An additional \(N=\) 26
were tested but removed from analyses because they had low scores on the
catch trials or because they were familiar with the non-English sound
stimuli we used in the adult experiment.

\subsubsection{Stimuli and similarity
rating}\label{stimuli-and-similarity-rating}

The sound stimuli were generated using the MBROLA Speech Synthesizer
(Dutoit, Pagel, Pierret, Bataille, \& Van der Vrecken, 1996). We
generated three kinds of nonsense word pairs which varied in their
degree of perceptual similarity to English speakers: 1) \emph{different}
pairs: \enquote{lif}/\enquote{neem} and \enquote{zem}/\enquote{doof}, 2)
\emph{intermediate} pairs: \enquote{aka}/\enquote{ama} and
\enquote{ada}/\enquote{aba}, and 3) \emph{similar} non-English pairs:
\enquote{ada}/\enquote{a\textipa{d\super h}a} (in hindi) and
\enquote{a\textipa{Q}a}/\enquote{a\textipa{\textcrh}a} (in arabic).

As for the objects, we used the Dynamic Stimuli javascript
library\footnote{https://github.com/erindb/stimuli} which allowed us to
generate objects in four different categories: \enquote{tree,}
\enquote{bird,} \enquote{bug,} and \enquote{fish.} These categories are
supposed to be naturally occurring kinds that might be seen on an alien
planet. In each category, we generated \emph{different},
\emph{intermediate} and \emph{similar} pairs by manipulating a
continuous property controlling features of the category's shape (e.g,
body stretch or head fatness).

In order to validate and quantify our similarity scales, we ran a
separate survey on Amazon Mechanical Turk where we asked \(N=20\) adults
participants to evaluate the similarity of each sound and object pair on
a 7-point scale. Data are shown in Figure \ref{fig:stim} where we scaled
responses within the range {[}0,1{]} for each stimulus group. These data
will be used in all models as a proxy for the perceptual distance
between the sound pairs (\(\Delta s\)) and the object pairs
(\(\Delta o\)).

\subsubsection{Design}\label{design}

Each age group saw only two of the three levels of similarity described
in the previous sub-section: \emph{different} vs. \emph{intermediate}
for the preschoolers, and \emph{intermediate} vs. \emph{similar} for
adults. We made this choice in light of pilot studies showing that
adults were at ceiling with \emph{different} sounds/objects, and
children were at chance with the \emph{similar} sounds/objects. That
said, this difference in the level of similarity is accounted for in the
model through using the appropriate distance (Figure \ref{fig:stim}).

\begin{figure}[h]

{\centering \includegraphics{ms_files/figure-latex/stim-1}

}

\caption{Distances for both sound and object pairs from an adult norming study. Data represent Likert values normalized to [0,1] interval. Error bars represent 95\% confidence intervals.}\label{fig:stim}
\end{figure}

To maximize our ability to measure subtle stimulus effects, the
experiment was a 2x2 within-subjects factorial design with four
conditions: high/low sound similarity crossed with high/low visual
object similarity. Besides the four conditions, we also tested
participants on a fifth catch condition which was similar in its
structure to the other ones but was used only to select participants who
were able to follow the instructions and show minimal learning.

\subsubsection{Procedure}\label{procedure}

Preschoolers were tested at the nursery school using a tablet, whereas
adults used their own computers to complete the same experiment online.
Participants were tested in a sequence of five conditions: the four
experimental conditions plus the catch condition. In each condition,
participants saw a first block of four exposure trials followed by four
testing trials, and a second block of two exposure trials (for memory
refreshment) followed by an additional four testing trials. The length
of this procedure was demanding, especially for children, but we adopted
a fully within-subjects design based on pilot testing that indicated
that precision of measurement was critical for testing our experimental
predictions.

In the exposure trials, participants saw two objects associated with
their corresponding sounds. We presented the first object on the left
side of the tablet's screen simultaneously with the corresponding sound.
The second sound-object association followed on the other side of the
screen after 500ms. For both objects, visual stimuli were present for
the duration of the sound clip (about 800ms). In the testing trials,
participants saw both objects simultaneously and heard only one sound.
They completed the trial by selecting which of the two objects
corresponded to the sound. The object-sound pairings were randomized
across participants, as was the order of the conditions (except for the
catch condition which was always placed in the middle of the testing
sequence). We also randomized the on-screen position (left vs.~right) of
the two pictures on each testing trial.

\subsection{Results}\label{results}

\begin{figure}[h]

{\centering \includegraphics{ms_files/figure-latex/allData-1}

}

\caption{Accuracy of word recognition as a function of the sound distance, the object distance, and the age group (preschool children vs. adults). We show both the models' predictions (dashed lines) and the experimental results (solid lines). The single-variance model uses one joint fitting parameter for both sound and meaning variances. The double-variance model uses two separate fittings parameters for the sound and the meaning variances. Error bars represent 95\% confidence intervals.}\label{fig:allData}
\end{figure}

Experimental results are shown in Figure \ref{fig:allData} (solid
lines). We first analyzed the results using a mixed-effects logistic
regression with sound distance, object distance and age group as fixed
effects, and with a maximal random effects structure (allowing us to
take into account the full nested structure of our data) (Barr, Levy,
Scheepers, \& Tily, 2013). We found main effects for all the fixed
effects in the regression. For the sound distance, we obtained
\(\beta =\) 0.52 (\(p\) \textless{} 0.001), replicating previous
findings that sound distance modulates success in word learning (e.g.,
Stager \& Werker, 1997). For object distance, we found \(\beta =\) 0.83
(\(p\) \textless{} 0.001), and this finding confirms the new prediction
of our model according to which object distance also modulates success
in word learning. Finally, for the age group, we obtained \(\beta =\)
0.76 (\(p\) \textless{} 0.001), showing that overall performance
improves with age.

\begin{table}

\caption{\label{tab:models}Characteristics and performance of the models used in this study.}
\centering
\begin{tabular}[t]{>{\bfseries}llrrrlrl}
\toprule
\multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{2}{c}{Children} & \multicolumn{2}{c}{Adults} \\
\cmidrule(l{2pt}r{2pt}){5-6} \cmidrule(l{2pt}r{2pt}){7-8}
Model & Structure & Param. & R\textsuperscript{2} & Sig\textsubscript{L} & Sig\textsubscript{C} & Sig\textsubscript{L} & Sig\textsubscript{C}\\
\midrule
model 1 & Sig\textsubscript{L} only & 1 & 0.27 & 1.00 & -- & 0.37 & --\\
model 2 & Sig\textsubscript{L} = Sig\textsubscript{C} & 1 & 0.95 & 0.60 & 0.6 & 0.15 & 0.15\\
model 3 & Sig\textsubscript{L} != Sig\textsubscript{C} & 2 & 1.00 & 0.83 & 0.31 & 0.12 & 0.17\\
\bottomrule
\end{tabular}
\end{table}

We next fit the three models obtained through expressions (1), (2), and
(3) to the participants' responses in each age group. The predictions of
the models are shown \ref{fig:allData} (dashed lines) and the parameter
estimates (for \(\sigma_L\) and \(\sigma_C\)) as well as models'
goodness to fit (i.e., measured through \(R^2\)) are presented in Table
\ref{tab:models}.

Model 1, which does not take into account ambiguity in recovering
concepts from observed objects, explains only a small part of the
variance. In contrast, Model 3, which does take into account this
ambiguity, accounts for all the variance. Interestingly, Model 2 which
has a single, shared uncertainty parameter for both auditory and visual
modalities still explains almost all the variance in human data,
suggesting that the explanatory power of the general Model 3 is largely
due to its structure, rather than its degrees of freedom.

As predicted, the uncertainty parameters were larger for children than
they were for adults (Table \ref{tab:models}), showing that the
probabilistic representations become more refined (that is, \(\sigma\)
becomes smaller) across development. Further, the parameter estimates of
Model 3 show that this developmental effect is more important for the
label-specific uncertainty that it is for the concept-specific
uncertainty.

\section{General Discussion}\label{general-discussion}

This paper explored the idea that some seemingly stage-like patterns in
cognitive development can be characterized in a continuous fashion. We
used as a case study the seminal work of Stager and Werker (1997)
showing a discrepancy between children's speech perception abilities and
their word learning skills. While this fact might seem like a specific
stage in early development, our model demonstrated, instead, that it can
be more simply understood in terms of continuous developmental change.

The main assumption of the model was that both word form and referent
are encoded in a probabilistic fashion. The model provided a
quantitative instantiation of the continuous development hypothesis
(Pajak et al., 2016; e.g., Swingley, 2007; Yoshida et al., 2009)). More
precisely, we showed that developmental changes in word-object mappings
can be characterized as a continuous refinement in the precision of the
probabilistic representations.

We find in the literature two broad accounts of development in the
Switch task: One that suggests \emph{direct} development of the sound
representation and one that hypothesizes \emph{indirect} development of
this representation through improvement in general cognitive resources.
On the first account, the sound representation becomes more precise as
learners refine the boundaries of their initially ambiguous phonetic
categories and as they gain more experience with the functional role of
these categories, i.e., contrasting word meaning (Apfelbaum \& McMurray,
2011; Dietrich, Swingley, \& Werker, 2007; Rost \& McMurray, 2009, 2010;
Yoshida et al., 2009). On the second account, the precision of sound
encoding in the switch task improves as a result of the maturation of
more general resources like the attentional and working memory capacity
(Hofer \& Levy, 2017; Stager \& Werker, 1997; Werker \& Fennell, 2004).
Such improvement allows older children and adults to better encode the
sound details while simultaneously matching these sounds to visual
objects. These two accounts are complementary and both seem to play a
role (see Tsui, Byers-Heinlein, \& Fennell, 2019 for a review).

Our model is compatible with both of these accounts. In our work, the
probability distributions do not distinguish between the direct and
indirect sources of uncertainty --- both are included. Indeed, part of
the measured uncertainty reflects the learner's degrees of confidence in
the phonetic/phonological boundaries (i.e., the direct account) and
another part reflects a possible drop in perceptual acuity due to high
cognitive load (i.e., the indirect account). Note, however, that the
model (at least in its current format) is incapable of answering
questions about the development of each of these sources of uncertainty
separately or about their relative contribution to the global
uncertainty.

Werker and Curtin (2005) proposed to explain development in the Switch
task using their theory called Processing Rich Information from
Multidimensional Interactive Representations (or PRIMIR) which purports
to explain various phenomena in early speech perception and word
learning within a unified framework. PRIMIR posits that children
initially try to attend to various features of the speech signal,
regardless of whether or not these features are relevant to the task at
hand. For example, when learning the meaning of similar sounds, infants
are unsure what detail is most important to identify words (i.e., the
phonemes), and will instead activate several aspects of the information
simultaneously (including, for example, the gender of the speaker). The
lack of attentional selection leads to confusion and then to failure in
the task.

According to PRIMIR, learning similar-sounding words becomes more robust
over time as children develop abstract phonemic categories. The latter
act as filters, allowing children to attend selectively to the important
information. This account is also compatible with our model (as it
resembles the direct account we discussed above). Developing a better
attentional strategy (thanks to phoneme) can translate into a reduction
in the uncertainty about whether a sound contrast signals a change in
meaning.

While most research focused on the sound representation in analyzing the
Switch task, this work showed that the visual representation of the
referent is equally important. Indeed, Model 1 --- which assumes that
any visually discriminable contrast can be encoded unambiguously as
separate referents --- failed to explain the data, whereas Model 2 and 3
--- which take into account visual ambiguity --- succeeded. As a
consequence of this assumption, we found that just like word learning is
modulated by the phonological similarity of the form, it is also
modulated by the visual similarity of the semantic referents.

Model 2, which predicts that sound similarity and visual similarity
influence word learning accuracy in the same way, explained slightly
less variance than Model 3 which predicts that these modalities
influence word learning differently. More precisely, a comparison of the
variance estimates across age groups shows that uncertainty reduction in
the visual modality was lower compared to that of the auditory modality
(Table 1). Perhaps this difference is due to the fact that, in our task,
the auditory speech had more sources of noise --- that children have to
deal with --- than the visual input does. The processing of speech
involved dealing with both perceptual noise and categorical ambiguity
(due to the fact that the phonemic boundaries are still developing). In
contrast, the processing of the visual input in our task involved only
perceptual noise and no category-related uncertainty.

Our finding that word learning is mediated by the visual similaérity of
the semantic objects has implications for theories of lexical
development. It suggests that, all things being equal, children may
prioritize the acquisition of words whose semantic referents are
visually different, as this allows them to minimize semantic ambiguity.
It will be interesting for future work to explore whether the results
that we obtained using visual similarity generalize to richer, more
conceptual features in the semantic space. This suggestion is, indeed,
supported by recent work investigating vocabulary development
(Engelthaler \& Hills, 2017; Fourtassi, Bian, \& Frank, 2018; Sizemore,
Karuza, Giusti, \& Bassett, 2018)

There are a few limitations to this work. One is that the model was fit
to data from children at a relatively older age (4-5 years old) than
what is typically studied in the literature (14-17 month-old). We
selected this older age group to optimize the number and precision of
the experimental measures (both are crucial to model fitting). Data
collection involved presenting participants with several trials across
four conditions in a between-subject design. It would have been
challenging to obtain such measures with infants. That said, though we
used data from older children, we still found clear differences compared
to adults, suggesting that development does not stop at 17 months, but
continues throughout childhood.

One limitation of the model is that it only accounts for bottom-up,
similarity-based effects. It does not account for how high-level factors
such as social and communicative cues can influence learning. For
example, Fennell and Waxman (2010) highlighted the fact that the Switch
task introduces novel words in isolation (e.g., \enquote{neem!}) rather
than within a naming phrase (e.g., \enquote{look at the neem!}). This
fact may prompt children to interpret these novel words in a
non-referential way (e.g., an exclamation such as \enquote{Wow!}).

To conclude, this paper proposes a model that accounts for the
development of an important aspect of word learning. Our account
suggests that the developmental data can be explained based on a
continuous process operating over similar representations across the
lifespan, suggesting developmental continuity. We used a case from word
learning as an example, but the same idea might apply to other aspects
of cognitive development that are typically thought of as stage-like
(e.g., acquisition of a theory of mind). Computational models, such as
the one proposed here, can help us investigate the extent to which such
discontinuities emerge due to genuine qualitative changes and the extent
to which they reflect the granularity of the researchers' own
measurement tools.

\vspace{1em}

\fbox{\parbox[b][][c]{14cm}{\centering All data and code for these analyses are available at\ \url{https://github.com/afourtassi/networks}}}
\vspace{1em}

\section{Acknowledgements}\label{acknowledgements}

This work was supported by a post-doctoral grant from the Fyssen
Foundation, NSF \#1528526, and NSF \#1659585.

\section{Disclosure statement}\label{disclosure-statement}

None of the authors have any financial interest or a conflict of
interest regarding this work and this submission.

\section{References}\label{references}

\setlength{\parindent}{-0.5in} \setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\hypertarget{ref-apfelbaum2011}{}
Apfelbaum, K. S., \& McMurray, B. (2011). Using variability to guide
dimensional weighting: Associative mechanisms in early word learning.
\emph{Cognitive Science}, \emph{35}.

\hypertarget{ref-barr2013}{}
Barr, D., Levy, R., Scheepers, C., \& Tily, H. (2013). Random effects
structure for confirmatory hypothesis testing: Keep it maximal.
\emph{Journal of Memory and Language}, \emph{68}(3).

\hypertarget{ref-carey2009}{}
Carey, S. (2009). \emph{The origin of concepts}. Oxford University
Press.

\hypertarget{ref-dautriche2015}{}
Dautriche, I., Swingley, D., \& Christophe, A. (2015). Learning novel
phonological neighbors: Syntactic category matters. \emph{Cognition},
\emph{143}.

\hypertarget{ref-dietrich2007}{}
Dietrich, C., Swingley, D., \& Werker, J. (2007). Native language
governs interpretation of salient speech sound differences at 18 months.
\emph{Proceedings of the National Academy of Sciences}, \emph{104}.

\hypertarget{ref-dutoit1996}{}
Dutoit, T., Pagel, V., Pierret, N., Bataille, F., \& Van der Vrecken, O.
(1996). The mbrola project: Towards a set of high quality speech
synthesizers free of use for non commercial purposes. In
\emph{Proceedings of ICSLP} (Vol. 3). IEEE.

\hypertarget{ref-engelthaler2017}{}
Engelthaler, T., \& Hills, T. T. (2017). Feature biases in early word
learning: Network distinctiveness predicts age of acquisition.
\emph{Cognitive Science}, \emph{41}.

\hypertarget{ref-fennell2010}{}
Fennell, C., \& Waxman, S. (2010). What paradox? Referential cues allow
for infant use of phonetic detail in word learning. \emph{Child
Development}, \emph{81}.

\hypertarget{ref-fernald2006}{}
Fernald, A., Perfors, A., \& Marchman, V. A. (2006). Picking up speed in
understanding: Speech processing efficiency and vocabulary growth across
the 2nd year. \emph{Developmental Psychology}, \emph{42}.

\hypertarget{ref-fourtassi2018}{}
Fourtassi, A., Bian, Y., \& Frank, M. C. (2018). Word learning as
network growth: A cross-linguistic analysis. In \emph{CogSci}.

\hypertarget{ref-golinkoff1987}{}
Golinkoff, R. M., Hirsh-Pasek, K., Cauley, K. M., \& Gordon, L. (1987).
The eyes have it: Lexical and syntactic comprehension in a new paradigm.
\emph{Journal of Child Language}, \emph{14}.

\hypertarget{ref-hofer2017}{}
Hofer, M., \& Levy, R. (2017). Modeling Sources of Uncertainty in Spoken
Word Learning. In \emph{Proceedings of the 39th Annual Meeting of the
Cognitive Science Society}.

\hypertarget{ref-lewis2013}{}
Lewis, M., \& Frank, M. (2013). An integrated model of concept learning
and word-concept mapping. In \emph{Proceedings of the annual meeting of
the cognitive science society} (Vol. 35).

\hypertarget{ref-pajak2016}{}
Pajak, B., Creel, S., \& Levy, R. (2016). Difficulty in learning
similar-sounding words: A developmental stage or a general property of
learning? \emph{Journal of Experimental Psychology: Learning, Memory,
and Cognition}, \emph{42}(9).

\hypertarget{ref-pater2004}{}
Pater, J., Stager, C., \& Werker, J. (2004). The perceptual acquisition
of phonological contrasts. \emph{Language}, \emph{80}.

\hypertarget{ref-piaget1954}{}
Piaget, J. (1954). \emph{The construction of reality in the child}. New
York, NY, US: Basic Books.

\hypertarget{ref-rost2009}{}
Rost, G., \& McMurray, B. (2009). Speaker variability augments
phonological processing in early word learning. \emph{Developmental
Science}, \emph{12}.

\hypertarget{ref-rost2010}{}
Rost, G., \& McMurray, B. (2010). Finding the signal by adding noise:
The role of noncontrastive phonetic variability in early word learning.
\emph{Infancy}, \emph{15}.

\hypertarget{ref-sizemore2018}{}
Sizemore, A. E., Karuza, E. A., Giusti, C., \& Bassett, D. S. (2018).
Knowledge gaps in the early growth of semantic feature networks.
\emph{Nature Human Behaviour}, \emph{2}(9).

\hypertarget{ref-stager1997}{}
Stager, C., \& Werker, J. (1997). Infants listen for more phonetic
detail in speech perception than in word-learning tasks. \emph{Nature},
\emph{388}(6640).

\hypertarget{ref-swingley2007}{}
Swingley, D. (2007). Lexical exposure and word-form encoding in
1.5-year-olds. \emph{Developmental Psychology}, \emph{43}(2).

\hypertarget{ref-thiessen2007}{}
Thiessen, E. (2007). The effect of distributional information on
children's use of phonemic contrasts. \emph{Journal of Memory and
Language}, \emph{56}.

\hypertarget{ref-tsui2019}{}
Tsui, A. S. M., Byers-Heinlein, K., \& Fennell, C. (2019). Associative
word learning in infancy: A meta-analysis of the switch task.
\emph{Developmental Psychology}, \emph{55}.

\hypertarget{ref-werker2005}{}
Werker, J., \& Curtin, S. (2005). PRIMIR: A developmental framework of
infant speech processing. \emph{Language Learning and Development},
\emph{1}.

\hypertarget{ref-werker2004}{}
Werker, J., \& Fennell, C. (2004). Listening to sounds versus listening
to words: Early steps in word learning. In D. G. Hall \& S. Waxman
(Eds.), \emph{Weaving a lexicon}. Cambridge: MIT Press.

\hypertarget{ref-werker1998}{}
Werker, J., Cohen, L. B., Lloyd, V. L., Casasola, M., \& Stager, C.
(1998). Acquisition of word--object associations by 14-month-old
infants. \emph{Developmental Psychology}, \emph{34}.

\hypertarget{ref-werker2002}{}
Werker, J., Fennell, C., Corcoran, K., \& Stager, C. (2002). Infants'
ability to learn phonetically similar words: Effects of age and
vocabulary size. \emph{Infancy}, \emph{3}.

\hypertarget{ref-white2013}{}
White, K., Yee, E., Blumstein, S., \& Morgan, J. (2013). Adults show
less sensitivity to phonetic detail in unfamiliar words, too.
\emph{Journal of Memory and Language}, \emph{68}(4).

\hypertarget{ref-yeung09}{}
Yeung, H., \& Werker, J. (2009). Learning words' sounds before learning
how words sound: 9-month-olds use distinct objects as cues to categorize
speech information. \emph{Cognition}, \emph{113}.

\hypertarget{ref-yoshida2009}{}
Yoshida, K., Fennell, C., Swingley, D., \& Werker, J. (2009).
14-month-olds learn similar-sounding words. \emph{Developmental
Science}, \emph{12}.

\end{document}